---
title: 'Risk of peritonitis in Malaysian incident PD patients: a competing risks approach'
author: "Mak Wen Yao, Ong Loke Meng"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  word_document: default
fontsize: 12pt
header-includes:
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{xcolor}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage[normalem]{ulem}
mainfont: Arial
classoption: a4paper
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r cleaning,eval=T,echo=F, cache=T, results=F, warning=F, message=F}
library(tidyverse)
# | Phase 1 competing risk 
sh1_patient <- read_csv("Sh1_Patient.csv")
sh3_AR <- read_csv("Sh2_PtAnnualAR.csv")
sh4_notif <- read_csv("Sh4_PtNotif.csv")
sh5_comorb <- read_csv("Sh5_PtNotifComorb.csv")
sh6_lab <- read_csv("Sh6_PtNotifLab.csv")
sh7_outcome <- read_csv("Sh7_PtNotifOutcome.csv")
sh8_peri <- read_csv("Sh8_PtNotifPDPeri.csv")

sh9_periAB <- read_csv("Sh9_PtNotifPDPeriAB.csv")
sh10_pdperiperf <- read_csv("Sh10_PtNotifPDPeriPerf.csv")
sh11_pdsys <- read_csv("Sh11_PtNotifPDSystem.csv")
sh12_sero <- read_csv("Sh12_PtNotifSerology.csv")
sh13_sdp <- read_csv("Sh13_SDP.csv")

```

```{r sh1,eval=T,echo=T, cache=T, results=T, warning=F, message=F}
#Step 1: clarify the data structure 
str(sh1_patient) #Date format ok, others should be factor 

x<-c("PatientID","PtSex","PtCitizenship","PtRace","PtABO") 
sh1_patient[,x]<-lapply(sh1_patient[,x],factor) #use lapply() to iteratively convert to factor variables 

#Step 2: identify duplicate patients 
which(duplicated(sh1_patient$PatientID)) #all unique patients 

#Step 3: identify ANY missing data 
which(is.na(sh1_patient)) #no missing data

#Step 4: Explore & enrich the data 

### 4.1: histogram for Race, and Blood-group 
library(DataExplorer)
plot_bar(sh1_patient$PtSex)         #Male=1, Female=2
plot_bar(sh1_patient$PtCitizenship) #Malaysian=1
plot_bar(sh1_patient$PtRace)        #Malay=1, Chi=2, Ind=3, BumiSabah=4
                                    #BumiSarawak=5, OrgAsliSemj=7, OtherMsian=19, 
                                    #Foreigner=20
plot_bar(sh1_patient$PtABO)         #A=1, B=2, AB=3, O=4, NA=8888, missing=9999
#note: some patients blood-group (PtABO) is "not available"; variable will not be used.

### 4.2: convert dob to age ###
sh1_patient['Age']<-NA #create a new variable called "Age"
sh1_patient['givendate']<-as.Date("31-12-2018", format="%d-%m-%Y")

library(lubridate) #use the lubridate package for dates 
sh1_patient$Age<- interval(start=sh1_patient$PtDateBirth, end=sh1_patient$givendate)/duration(num = 1, units = "years") 

ggplot(sh1_patient, aes(x=Age)) + 
  geom_histogram()                  #histogram shows negative age 

#   |to identify which patients had negative age 
subset(sh1_patient, sh1_patient$givendate < sh1_patient$PtDateBirth) #70695,71570
sh1_patient$PtDateBirth[which(sh1_patient$PatientID==70695)]<- as.Date("1929-07-16") #no need to specify "format", or else will turn into NA 
sh1_patient$PtDateBirth[which(sh1_patient$PatientID==71570)]<- as.Date("1929-09-04")

#   |re-run age calculation 
sh1_patient$Age<- interval(start=sh1_patient$PtDateBirth, end=sh1_patient$givendate)/duration(num = 1, units = "years") 

ggplot(sh1_patient, aes(x=Age)) + 
  geom_histogram()                  #skewness to the left 

```

```{r sh3,eval=T,echo=T, cache=T, results=T, warning=F, message=F}

#Step 1: clarify the data structure 
str(sh3_AR) #PtARID, PatientID, NotifID, SDPID, PtAssist should be factor 
x<-c("PtARID", "PatientID", "NotifID","PtRRTID", "SDPID")
sh3_AR[,x]<-lapply(sh3_AR[,x],factor)

#Step 2: identify duplicate patients 
which(duplicated(sh3_AR$PatientID)) #show multiple entries from same patients 

#Step 3: identify ANY missing data 
which(is.na(sh3_AR))                #no missing data 
 
#Step 4: Explore & enrich the data 

## 4.1: arrange row according to PatientID then DataYear 
sh3_AR<-arrange(sh3_AR,PatientID, DataYear) #some patients did not return the following year -> died? lost to f/up?

## 4.2: identify which PtAssist changed / deteriorates 
#plot_bar(sh3_AR$PtAssist)
which(sh3_AR$PtAssist==8888) #26 rows has NA for PtAssist 
sh3_AR$PtAssist[which(sh3_AR$PtAssist==8888)]<-NA
which(is.na(sh3_AR$PtAssist)) #all 26 rows converted to NA

## 4.2.1: PtAssist Changed from the year before 
sh3_AR<-sh3_AR %>%
  group_by(PatientID) %>%
  mutate(Asst.change=ifelse(PtAssist==lag(PtAssist), 0, 1) ) #lag() the previous row

sh3_AR$Asst.change[which(is.na(sh3_AR$Asst.change))]<-0

## 4.2.2: PtAssist deteriorated from the year before 
sh3_AR<-sh3_AR %>%
  group_by(PatientID) %>%
  mutate(deteriorate=ifelse(PtAssist>lag(PtAssist), 1, 0)) #lag() the previous row

sh3_AR$deteriorate[which(is.na(sh3_AR$deteriorate))]<-0


## Create a new sub-data that contains latest DataYear
## Will sacrifice some details in PtAssist over time 
sh3_AR<-arrange(sh3_AR, PatientID, desc(DataYear))

library(data.table)
setDT(sh3_AR)
sh3_AR2<-sh3_AR[rowid(rleid(PatientID)) == 1]
sh3_AR2<-subset(sh3_AR2, select =  -c(8:9))

``` 

```{r sh4,eval=T,echo=T, cache=T, results=T, warning=F, message=F}

#Step 1: clarify the data structure 
str(sh4_notif)    #A few issues: 
                  #PatientID, NotifID, PtRRTID, SDPID - factor
                  #PtDateCommence, PtDateCentre, ESRFDateFirstRx - date 
                  #PtAssis - no.obs less than sh3_AR; ignore PtAssist in sh4_notif 
                  #Modality in 'character' but all should be PD pt; ignore 

x<-c("PatientID", "NotifID","PtRRTID", "SDPID")
sh4_notif[,x]<-lapply(sh4_notif[,x],factor)

func<- function(x) as.Date(x,format = "%d-%m-%y") #create a function to convert to Date
sh4_notif[,6:8]<-data.frame(lapply(sh4_notif[,6:8], func)) #lapply() to convert to Date

sh4_notif<-subset(sh4_notif, select= -c(15))

sh4_notif$PtRenal1<-as.factor(sh4_notif$PtRenal1)

#Step 2: identify duplicate patients 
sh4_notif<-arrange(sh4_notif, PatientID, PtDateCentre) #arrange by PtDateCentre; 
                                                       #repeated entry due to patient 
                                                       #changing treatment site 
which(duplicated(sh4_notif$PatientID))                 #69 repeat entries by PatientID 

## 2.1: subset patients who changed site -> ESRFDate not complete/missing 
sh4_subset<- sh4_notif %>%
  group_by(PatientID) %>%
  filter(n()!=1)               #this method produce 129 observations 
  
  
sh4_subset<-subset(sh4_notif, (duplicated(sh4_notif$PatientID) | duplicated(sh4_notif$PatientID, fromLast = TRUE)))     # this also produce 129 obs

#only one subject show different ESRFDateFirstRx; 66565 (SDPID 530)
sh4_notif$ESRFDateFirstRx[which(sh4_notif$PatientID==66565 & sh4_notif$SDPID==530)]<-as.Date("2013-03-20")

## Data exploration with sh4_subset 
##  | Aim to remove duplicate and see if any vital data gone missing 
sh4_subset$ESRFDateFirstRx[which(sh4_subset$PatientID==66565 & sh4_subset$SDPID==530)]<-as.Date("2013-03-20")
sh4_subset<-arrange(sh4_subset, PatientID, ESRFDateFirstRx)

##  | Use rleid() to generate run-length type group ID 
setDT(sh4_subset)
sh4_subset<-sh4_subset[rowid(rleid(PatientID)) == 1]

#75148, 75322, 75339, 75781, 75782: 5 pt missing PtRenal1
#double-check original dataset, confirm all 5 data fields missing

#before we make any changes to ESRFDateFirstRx, arrange() the dataset by PatientID, ESRFDateFirstRx
sh4_notif<-sh4_notif %>%
  group_by(PatientID) %>%
  arrange(sh4_notif, ESRFDateFirstRx, .by_group = TRUE)
#change ESRFDateFirstRx to character 
sh4_notif$ESRFDateFirstRx<-as.character(sh4_notif$ESRFDateFirstRx)

sh4_notif<- sh4_notif %>% 
  group_by(PatientID) %>%
  fill(ESRFDateFirstRx)
#change ESRFDateFirstRx back to date 
sh4_notif$ESRFDateFirstRx<-as.Date(sh4_notif$ESRFDateFirstRx)

#na_bool = is.na(sh4_notif$ESRFDateFirstRx)
#sh4_notif[na_bool,]$ESRFDateFirstRx=sh4_notif[c(FALSE, na_bool[-length(na_bool)]),]$ESRFDateFirstRx #shift the value up if ESRFDateFirstRx is NA

## 2.2: PtRenal1 
plot_bar(sh4_notif$PtRenal1) #quite some patient w/out valid reasons 

#Step 3: identify ANY missing data 
which(is.na(sh4_notif[,8])) #index 668 & 1217 have real missing data; 
                            #PatientID:72535,1003737
                            #should we assume that ESRF date starts with Dialysis date? Yes 

sh4_notif$ESRFDateFirstRx[which(sh4_notif$PatientID==72535)]<-as.Date("2014-10-12")
sh4_notif$ESRFDateFirstRx[which(sh4_notif$PatientID==1003737)]<-as.Date("2014-12-21")
                            

#Step 4: Explore & enrich the data 

## 4.1: create two new variables "times since diagnosis" and "time since 1st PD"
sh4_notif['day_diag'] <- NA
sh4_notif['day_pd'] <- NA 
sh4_notif['givendate']<-NA 

sh4_notif$givendate <- as.Date("31-12-2018", format="%d-%m-%Y")
sh4_notif$day_pd <- sh4_notif$givendate - sh4_notif$PtDateCommence
sh4_notif$day_diag <- sh4_notif$givendate - sh4_notif$ESRFDateFirstRx #the two cases above will have time calculated from PtDateCommence  
                                                      
## 4.2: remove all the duplicates
setDT(sh4_notif)
sh4_notif<-sh4_notif[rowid(rleid(PatientID)) == 1] #final dataset has 1149 obs

``` 

```{r sh5,eval=T,echo=T, cache=T, results=T, warning=F, message=F}

#Step 1: clarify the data structure 
str(sh5_comorb) #Issues identified 
                #PtNotifComorbidityID, PatientID, NotifID, PtRRTID, SDPID, Comorbidity - factor 
                #Drop DateDiagnosis, DateDiagnosis_Est, DateDiagnosis_Unk, DateResolve - all missing 

x<-c("PtNotifComorbidityID", "PatientID", "NotifID", "PtRRTID", "SDPID", "Comorbidity")
sh5_comorb[,x]<-lapply(sh5_comorb[,x], factor)

sh5_comorb<-subset(sh5_comorb, select = -c(8:11)) 

#Step 2: identify duplicate patients 
which(duplicated(sh5_comorb$PatientID)) #PatientID repeated for multiple comorbidities 
                                        #Should convert long to wide, each comorb is a variable 
                                        #See Step 4 for details 

#Step 3: identify ANY missing data 
which(is.na(sh5_comorb[,1:8])) #no missing data

#Step 4: Explore & enrich the data 

## 4.1: convert the column "Comorbidity" from long to wide
## | use the newer pivot_wider() from tidyverse/tidyr to make the conversion 
## | add a new column to indicate comorb is "present" 
sh5_comorb['comob_present']<-as.integer(1)
class(sh5_comorb$comob_present)

sh5_comorb <- subset(sh5_comorb, select = c(2,8,10))
sh5_comorb<-as_tibble(sh5_comorb)

sh5_comorb<-  sh5_comorb %>% 
  pivot_wider(
    names_from = Comorbidity,
    values_from = comob_present)

    
    #    values_fill = as.list("0")    #must force the "0" as the "list" type, or else it wouldnt work
  

## |Comorbidity type (change column name)
## comob.hpt, comob.dm, comob.peptic_ulcer, comob.other_cardiac, comob.cerebrovas, comob.chro_res, comob.ihd, comob.other, comob.cancer, comob.ccf, comob.impair_vision, comob.amputation, comob.chron_liver, comob.peri_vas, comob.psy, comob.GI, comob.cvd, comob.tb, comob.ment_retard, comob.dementia, comob.sub_abuse, 

## rename all columns 

names(sh5_comorb)<- c("PatientID", "comob.hpt", "comob.dm", "comob.peptic_ulcer", "comob.other_cardiac", "comob.cerebrovas", "comob.chro_res", "comob.ihd", "comob.other", "comob.cancer", "comob.ccf", "comob.impair_vision", "comob.amputation", "comob.chron_liver", "comob.peri_vas", "comob.psy", "comob.GI", "comob.cvd", "comob.tb", "comob.ment_retard", "comob.dementia", "comob.sub_abuse")

sh5_comorb$comob.dm[which(sh5_comorb$comob.dm=="NULL")]<-"0"
sh5_comorb$comob.hpt[which(sh5_comorb$comob.hpt=="NULL")]<-"0"
sh5_comorb$comob.peptic_ulcer[which(sh5_comorb$comob.peptic_ulcer=="NULL")]<-"0"
sh5_comorb$comob.other_cardiac[which(sh5_comorb$comob.other_cardiac=="NULL")]<-"0"
sh5_comorb$comob.cerebrovas[which(sh5_comorb$comob.cerebrovas=="NULL")]<-"0"
sh5_comorb$comob.chro_res[which(sh5_comorb$comob.chro_res=="NULL")]<-"0"
sh5_comorb$comob.ihd[which(sh5_comorb$comob.ihd=="NULL")]<-"0"
sh5_comorb$comob.other[which(sh5_comorb$comob.other=="NULL")]<-"0"
sh5_comorb$comob.cancer[which(sh5_comorb$comob.cancer=="NULL")]<-"0"
sh5_comorb$comob.ccf[which(sh5_comorb$comob.ccf=="NULL")]<-"0"
sh5_comorb$comob.impair_vision[which(sh5_comorb$comob.impair_vision=="NULL")]<-"0"
sh5_comorb$comob.amputation[which(sh5_comorb$comob.amputation=="NULL")]<-"0"
sh5_comorb$comob.chron_liver[which(sh5_comorb$comob.chron_liver=="NULL")]<-"0"
sh5_comorb$comob.peri_vas[which(sh5_comorb$comob.peri_vas=="NULL")]<-"0"
sh5_comorb$comob.psy[which(sh5_comorb$comob.psy=="NULL")]<-"0"
sh5_comorb$comob.GI[which(sh5_comorb$comob.GI=="NULL")]<-"0"
sh5_comorb$comob.cvd[which(sh5_comorb$comob.cvd=="NULL")]<-"0"
sh5_comorb$comob.tb[which(sh5_comorb$comob.tb=="NULL")]<-"0"
sh5_comorb$comob.ment_retard[which(sh5_comorb$comob.ment_retard=="NULL")]<-"0"
sh5_comorb$comob.dementia[which(sh5_comorb$comob.dementia=="NULL")]<-"0"
sh5_comorb$comob.sub_abuse[which(sh5_comorb$comob.sub_abuse=="NULL")]<-"0"

sh5_comorb$PatientID<-as.character(sh5_comorb$PatientID)

sh5_comorb$comob.dm[which(sh5_comorb$comob.dm=="1")]<-1
sh5_comorb$comob.hpt[which(sh5_comorb$comob.hpt=="1")]<-1
sh5_comorb$comob.hpt[which(sh5_comorb$PatientID=="78556")]<-1
sh5_comorb$comob.peptic_ulcer[which(sh5_comorb$comob.peptic_ulcer=="1")]<-1
sh5_comorb$comob.other_cardiac[which(sh5_comorb$comob.other_cardiac=="1")]<-1
sh5_comorb$comob.cerebrovas[which(sh5_comorb$comob.cerebrovas=="1")]<-1
sh5_comorb$comob.chro_res[which(sh5_comorb$comob.chro_res=="1")]<-1
sh5_comorb$comob.ihd[which(sh5_comorb$comob.ihd=="1")]<-1
sh5_comorb$comob.other[which(sh5_comorb$comob.other=="1")]<-1
sh5_comorb$comob.cancer[which(sh5_comorb$comob.cancer=="1")]<-1
sh5_comorb$comob.ccf[which(sh5_comorb$comob.ccf=="1")]<-1
sh5_comorb$comob.impair_vision[which(sh5_comorb$comob.impair_vision=="1")]<-1
sh5_comorb$comob.amputation[which(sh5_comorb$comob.amputation=="1")]<-1
sh5_comorb$comob.chron_liver[which(sh5_comorb$comob.chron_liver=="1")]<-1
sh5_comorb$comob.peri_vas[which(sh5_comorb$comob.peri_vas=="1")]<-1
sh5_comorb$comob.psy[which(sh5_comorb$comob.psy=="1")]<-1
sh5_comorb$comob.GI[which(sh5_comorb$comob.GI=="1")]<-1
sh5_comorb$comob.cvd[which(sh5_comorb$comob.cvd=="1")]<-1
sh5_comorb$comob.tb[which(sh5_comorb$comob.tb=="1")]<-1
sh5_comorb$comob.ment_retard[which(sh5_comorb$comob.ment_retard=="1")]<-1
sh5_comorb$comob.dementia[which(sh5_comorb$comob.dementia=="1")]<-1
sh5_comorb$comob.sub_abuse[which(sh5_comorb$comob.sub_abuse=="1")]<-1


sh5_comorb$comob.dm<-as.numeric(sh5_comorb$comob.dm)
sh5_comorb$comob.hpt<-as.numeric(sh5_comorb$comob.hpt) #Id 78556
sh5_comorb$comob.peptic_ulcer<-as.numeric(sh5_comorb$comob.peptic_ulcer)
sh5_comorb$comob.other_cardiac<-as.numeric(sh5_comorb$comob.other_cardiac)
sh5_comorb$comob.cerebrovas<-as.numeric(sh5_comorb$comob.cerebrovas)
sh5_comorb$comob.chro_res<-as.numeric(sh5_comorb$comob.chro_res)
sh5_comorb$comob.ihd<-as.numeric(sh5_comorb$comob.ihd)
sh5_comorb$comob.other<-as.numeric(sh5_comorb$comob.other)
sh5_comorb$comob.cancer<-as.numeric(sh5_comorb$comob.cancer)
sh5_comorb$comob.ccf<-as.numeric(sh5_comorb$comob.ccf)
sh5_comorb$comob.impair_vision<-as.numeric(sh5_comorb$comob.impair_vision)
sh5_comorb$comob.amputation<-as.numeric(sh5_comorb$comob.amputation)
sh5_comorb$comob.chron_liver<-as.numeric(sh5_comorb$comob.chron_liver)
sh5_comorb$comob.peri_vas<-as.numeric(sh5_comorb$comob.peri_vas)
sh5_comorb$comob.psy<-as.numeric(sh5_comorb$comob.psy)
sh5_comorb$comob.GI<-as.numeric(sh5_comorb$comob.GI)
sh5_comorb$comob.cvd<-as.numeric(sh5_comorb$comob.cvd)
sh5_comorb$comob.tb<-as.numeric(sh5_comorb$comob.tb)
sh5_comorb$comob.ment_retard<-as.numeric(sh5_comorb$comob.ment_retard)
sh5_comorb$comob.dementia<-as.numeric(sh5_comorb$comob.dementia)
sh5_comorb$comob.sub_abuse<-as.numeric(sh5_comorb$comob.sub_abuse)
sh5_comorb$PatientID<-as.numeric(sh5_comorb$PatientID)



``` 

```{r sh7,eval=T,echo=T, cache=T, results=T, warning=F, message=F}

#Step 1: clarify the data structure 
str(sh7_outcome)  #Similar issues as previous sheets 
                  #PtNotifOutcomeID, PatientID, NotifID, PtRRTID, SDPID - factor 
                  #Other outcome variables should be in "factor" format but leave them be numeric for now 

x<-c("PtNotifOutcomeID", "PatientID", "NotifID", "PtRRTID", "SDPID", "PtOutcomeID")
sh7_outcome[,x]<-lapply(sh7_outcome[,x], factor)

sh7_outcome$PtDateOutcome <- as.Date(sh7_outcome$PtDateOutcome, format = "%d-%m-%y")
sh7_outcome$PtDateOutcomeNotified <- as.Date(sh7_outcome$PtDateOutcomeNotified, format="%d-%m-%y")


#Step 2: identify duplicate patients 
sh7_outcome<-arrange(sh7_outcome, PatientID, PtDateOutcome)

which(duplicated(sh7_outcome$PatientID))  #identify 172 duplicated records - multiple outcomes by the same patient 

#Step 3: identify ANY missing data 

## 3.1: important variable is the "PtDateOutcome. Missing values in PtDateOutcomeNotified are ignored
which(is.na(sh7_outcome$PtDateOutcome))  #no missing data 


#Step 4: Explore & enrich the data 

## 4.1: check if "death" (terminal event) occurs befor any other outcome; death ==1 
sh7_outcome2<-subset(sh7_outcome, !sh7_outcome$PtOutcomeID==0)
sh7_outcome2<- sh7_outcome2[,1:9]

sh7_outcome3<- sh7_outcome2 %>%
  group_by(PatientID, desc(PtOutcomeID)) %>%
  filter(n()!=1) %>%
  mutate(wrong=ifelse((PtDateOutcome > lag(PtDateOutcome)),1,0)) #No event occurs after death; 

## 4.2: drop the column "modality" 
sh7_outcome<-subset(sh7_outcome, select =  -c(5))

## 4.3: drop the row with PtOutcomeID == 5 (move to other centre)
sh7_outcome<-subset(sh7_outcome, !(PtOutcomeID==5 | PtOutcomeID==0))
plot_bar(sh7_outcome$PtOutcomeID)

## 4.3: shorten the column 
sh7_outcome<-subset(sh7_outcome, select = c(1:9))
``` 

```{r sh8,eval=T,echo=T, cache=T, results=T, warning=F, message=F}

#Step 1: clarify the data structure 
str(sh8_peri)     #A few issues
                  #PtNotifPDPeriID, PatientID, NotifID, PtRRTID, SDPID, PtARID, PeritonitisType - Factor 
                  #DatePeri - Date

x<-c("PtNotifPDPeriID", "PatientID", "NotifID", "PtRRTID", "SDPID", "PtARID")
sh8_peri[,x]<-lapply(sh8_peri[,x], factor)

sh8_peri$DatePeri<-as.Date(sh8_peri$DatePeri, format="%d-%m-%y")

#Step 2: identify duplicate patients 
which(duplicated(sh8_peri$PatientID)) #541 repeated instances 

#Step 3: identify ANY missing data 
## | important variabls are DatePeri and PeritonitisType 
which(is.na(sh8_peri$DatePeri) | is.na(sh8_peri$PeritonitisType)) #no missing data

#Step 4: Explore & enrich the data 
## | arrange according to PatientID and DatePeri, and complete the remaining analysis in following sections

sh8_peri<-sh8_peri %>%
  arrange(PatientID, DatePeri)

sh8_peri<-subset(sh8_peri,select= c(1:9))


##  | some issues identified from preliminary analysis: 
##    1. Wrong PtDateOutcome (!) 70674, peri 2005-09-17;  75995 peri 2006-10-23; 
##    2. 70766, move to other centre, 2014-03-01; 
##    3. 72388, ESI 2014-09-01; 74135, peri 2014-05-19;76595 peri 2014-02-08 (date ESI is correct, need to check PtDateCommence)

# Step 1: Correct wrong data entry (PatientID == 70674, 75995)
sh8_peri$DatePeri[(sh8_peri$PatientID==70674 & sh8_peri$PtNotifPDPeriID==11619)]<-as.Date("2015-09-17")

sh8_peri$DatePeri[(sh8_peri$PatientID==75995 |sh8_peri$PtNotifPDPeriID==13695)]<-as.Date("2016-10-23") #this subject reported multipl ESI & peri on the same date 

# Step 2: Cannot verify date of peri, possibly experience before formally enrolled in dialysis program 
# to remove the incidents 
which(sh8_peri$PatientID==72388 & sh8_peri$PtNotifPDPeriID==10874) #index 612
sh8_peri<-sh8_peri[-612,]

which(sh8_peri$PatientID==74135 & sh8_peri$PtNotifPDPeriID==10309) #index 715
sh8_peri<-sh8_peri[-715,]

which(sh8_peri$PatientID==76595 & sh8_peri$PtNotifPDPeriID==11103) #index 1049
sh8_peri<-sh8_peri[-1049,]

# Step 3: Remove ESI from cases where ESI and Peritonitis co-exist on the SAME DATE 
sh8_peri2<- sh8_peri %>% 
  group_by(PatientID, DatePeri) %>%
  filter(n()!=1) %>%
  filter(DatePeri == lead(DatePeri))

x<-which(sh8_peri$DatePeri == lead(sh8_peri$DatePeri)) 

sh8_peri <- sh8_peri[-c(129,258,261,303,346,349,425,521,562,619,621,637,646,680,757,778,877,921,933,976,977,978,1031),]

# Double check if the removal is successful 
which(sh8_peri$DatePeri == lead(sh8_peri$DatePeri)) 

# Step 4: Many subjects have same date peritonitis with PtDateCommence 
#      4.1: Some subjects have same peritonitis date with death or transfer to HD 

#70408, 2014-05-12 +1
#70417, 2017-05-15 (same date as transfer to HD) -1
#70685, 2014-03-23  +1
#70687, 2014-04-27 +1
#70715, 2015-03-30 (same date as death) -1
#70854, 2017-06-17 (same date as death) -1
#71669, 2014-08-04 +1
#73942, 2014-08-04 +1
#74147, 2017-07-17 (same date as transfer to HD) -1
#76181, 2016-10-17 (same date as transfer to HD) -1
#76394, 2017-11-18 (same date as death) -1

sh8_peri$DatePeri[(sh8_peri$PatientID==70408 |sh8_peri$PtNotifPDPeriID==11016)]<-as.Date("2014-05-14")

sh8_peri$DatePeri[(sh8_peri$PatientID==70685 |sh8_peri$PtNotifPDPeriID==10254)]<-as.Date("2014-03-25")

sh8_peri$DatePeri[(sh8_peri$PatientID==70687 |sh8_peri$PtNotifPDPeriID==10260)]<-as.Date("2014-04-29")

sh8_peri$DatePeri[(sh8_peri$PatientID==71669 |sh8_peri$PtNotifPDPeriID==10705)]<-as.Date("2014-08-06")

sh8_peri$DatePeri[(sh8_peri$PatientID==73942 |sh8_peri$PtNotifPDPeriID==10250)]<-as.Date("2014-08-06")

sh8_peri$DatePeri[(sh8_peri$PatientID==70417 |sh8_peri$PtNotifPDPeriID==16027)]<-as.Date("2017-05-13")

sh8_peri$DatePeri[(sh8_peri$PatientID==70715 |sh8_peri$PtNotifPDPeriID==12029)]<-as.Date("2015-03-28")

sh8_peri$DatePeri[(sh8_peri$PatientID==70854 |sh8_peri$PtNotifPDPeriID==14790)]<-as.Date("2017-06-15")

sh8_peri$DatePeri[(sh8_peri$PatientID==74147 |sh8_peri$PtNotifPDPeriID==15678)]<-as.Date("2017-07-15")

sh8_peri$DatePeri[(sh8_peri$PatientID==76181 |sh8_peri$PtNotifPDPeriID==13873)]<-as.Date("2016-10-16")

sh8_peri$DatePeri[(sh8_peri$PatientID==76394 |sh8_peri$PtNotifPDPeriID==15339)]<-as.Date("2017-11-16")

## Step 5: Some patients have double entry 
#rep<-sh8_peri %>%
#  group_by(PatientID, PtDateOutcome) %>%
#  filter(n()!=1) %>% #a total of 5 unique subject id with 18 obs 
#  filter(PtDateOutcome==lead(PtDateOutcome)) #this filters down to 13 obs (correctly minus 5)




which(sh8_peri$PatientID==70685 & sh8_peri$PtNotifPDPeriID==10254)
which(sh8_peri$PatientID==71669 & sh8_peri$PtNotifPDPeriID==10705)
which(sh8_peri$PatientID==71669 & sh8_peri$PtNotifPDPeriID==13614)
which(sh8_peri$PatientID==71669 & sh8_peri$PtNotifPDPeriID==13615)
which(sh8_peri$PatientID==71669 & sh8_peri$PtNotifPDPeriID==13616)
which(sh8_peri$PatientID==71669 & sh8_peri$PtNotifPDPeriID==13617)
which(sh8_peri$PatientID==71669 & sh8_peri$PtNotifPDPeriID==13618)
which(sh8_peri$PatientID==71669 & sh8_peri$PtNotifPDPeriID==13619)
which(sh8_peri$PatientID==71669 & sh8_peri$PtNotifPDPeriID==13620)
which(sh8_peri$PatientID==73942 & sh8_peri$PtNotifPDPeriID==10250)
which(sh8_peri$PatientID==76181 & sh8_peri$PtNotifPDPeriID==12162)
which(sh8_peri$PatientID==76181 & sh8_peri$PtNotifPDPeriID==13872)
which(sh8_peri$PatientID==76394 & sh8_peri$PtNotifPDPeriID==15337)


#index 394,531,532,533,534,535,536,537,538,673,992,993,1016
sh8_peri <- sh8_peri[-c(394,531,532,533,534,535,536,537,538,673,992,993,1016),]
  
sh8_peri <- subset(sh8_peri, sh8_peri$PeritonitisType==2 )

``` 

```{r sh7_8, eval=T, echo=F, cache=T, results=FALSE, warning=FALSE, message=FALSE} 

# combine both sh7 and sh8 into one outcome dataset 

## Step 1: change the PeritonitisType to PtOutcomeID, and DatePeri to PtDateOutcome 
##  | Peritonitis from 2-> 12; ESI from 1-> 13 
sh8_peri$PeritonitisType[sh8_peri$PeritonitisType==2]<- 12
sh8_peri$PeritonitisType[sh8_peri$PeritonitisType==1]<- 13

##  | Rename column 
sh8_peri<- rename(sh8_peri, PtOutcomeID=PeritonitisType)
sh8_peri<- rename(sh8_peri, PtDateOutcome=DatePeri)

##  | rbind both dataset with data.table()
library(data.table)
z = list(sh7_outcome, sh8_peri)
sh7_8<- rbindlist(z, use.names = TRUE, fill = TRUE)

#sh7_8<-merge(sh7_outcome, sh8_peri, by="PatientID", all = TRUE)

## Step 2: re-arrange the new combined dataset with PatientID 
sh7_8<-arrange(sh7_8, PatientID, PtDateOutcome)

which(is.na(sh7_8$PtOutcomeID) | is.na(sh7_8$PtDateOutcome)) #no missing data 

## Step 3: create event; t.start, t.stop, time cannot create without PtOutcomeDate 

sh7_8<-sh7_8 %>%
  mutate(event=ifelse(PtOutcomeID==12, 1,0)) %>%
  relocate(event, .after=PtOutcomeID)

sh7_8 <- subset(sh7_8,select =  c(2:8))

```

```{r sh_demo, eval=T, echo=F, cache=T, results=FALSE, warning=FALSE, message=FALSE}

# combine sh1, sh3, sh4, and sh5 to form a combined demographic dataset 

library(data.table)
#combine sh1_patient and sh5_comorb first 
sh_demo1<-merge(sh1_patient, sh5_comorb, by="PatientID", all = TRUE)
sh_demo2<-merge(sh_demo1, sh4_notif, by="PatientID", all=TRUE)
sh_demo.fin<-merge(sh3_AR2,sh_demo2, by="PatientID", all = TRUE)

sh_demo.fin<-arrange(sh_demo.fin, PatientID) #1149 obs 

```

```{r sh_11, eval=T, echo=F, cache=T, results=FALSE, warning=FALSE, message=FALSE}

sh11_pdsys <- read_csv("Sh11_PtNotifPDSystem.csv")


#Step 1: clarify the data structure 
str(sh11_pdsys) #PDRegimen and PDSystem were both numeric variables 
                #DateStart and DateEnd were character variables 

x<-c("PDRegimen", "PDSystem", "PatientID")
sh11_pdsys[,x]<-lapply(sh11_pdsys[,x], factor)

sh11_pdsys$DateStart <- as.Date(sh11_pdsys$DateStart, format = "%d-%m-%y")
sh11_pdsys$DateEnd <- as.Date(sh11_pdsys$DateEnd, format = "%d-%m-%y")

#set PDRegimen ==8888 or 9999 to NA 
sh11_pdsys$PDRegimen[which(sh11_pdsys$PDRegimen==8888 | sh11_pdsys$PDRegimen==9999)]<-NA

sh11_pdsys <- sh11_pdsys %>%
  group_by(PatientID) %>%
  arrange(PatientID, DateStart) %>%
  mutate(Index=row_number()) %>%
  relocate(Index, .after=PDRegimen)

which(is.na(sh11_pdsys$PDRegimen)&sh11_pdsys$Index==1) #1 (ID:1253), 79 (ID: 25299), 1601 (ID:70770), 1754 (ID: 71009)

#change PDRegimen into a factor variable 
sh11_pdsys$PDRegimen<- as.factor(sh11_pdsys$PDRegimen)

sh11_pdsys$PDRegimen[which(sh11_pdsys$PatientID==1253&sh11_pdsys$Index==1)]<-3
sh11_pdsys$PDRegimen[which(sh11_pdsys$PatientID==70770&sh11_pdsys$Index==1)]<-1
sh11_pdsys$PDRegimen[which(sh11_pdsys$PatientID==71009&sh11_pdsys$Index==1)]<-1
sh11_pdsys$PDRegimen[which(sh11_pdsys$PatientID==25299&sh11_pdsys$Index==1)]<-1

#set PDSystem ==8888 or 9999 to NA 
sh11_pdsys$PDSystem[which(sh11_pdsys$PDSystem==8888 | sh11_pdsys$PDSystem==9999)]<-NA

sh11_pdsys<- relocate(sh11_pdsys, PDSystem, .after = Index)

which(is.na(sh11_pdsys$PDSystem)&sh11_pdsys$Index==1) #1 (ID:1253), 79 (ID: 25299), 1601 (ID:70770), 1754 (ID: 71009)

sh11_pdsys$PDSystem[which(sh11_pdsys$PatientID==1253&sh11_pdsys$Index==1)]<-101
sh11_pdsys$PDSystem[which(sh11_pdsys$PatientID==70770&sh11_pdsys$Index==1)]<-101
sh11_pdsys$PDSystem[which(sh11_pdsys$PatientID==71009&sh11_pdsys$Index==1)]<-101
sh11_pdsys$PDSystem[which(sh11_pdsys$PatientID==25299&sh11_pdsys$Index==1)]<-101

## Keep only the first row 
sh11_pdsys <- sh11_pdsys %>%
  arrange(PatientID, Index) %>%
  filter(Index==1)

which(is.na(sh11_pdsys$PDRegimen) | is.na(sh11_pdsys$PDSystem)) #no missing data 

## trim the dataset to contain only PatientID, PDSystem, and PDRegimen 
sh11_pdsys<- select(sh11_pdsys, c("PatientID", "PDSystem", "PDRegimen"))

## create new variable for PDSystem as "brand" 
sh11_pdsys['brand']<-NA
sh11_pdsys$brand[which(sh11_pdsys$PDSystem==101|sh11_pdsys$PDSystem==102|sh11_pdsys$PDSystem==103|sh11_pdsys$PDSystem==104|sh11_pdsys$PDSystem==105|sh11_pdsys$PDSystem==106|sh11_pdsys$PDSystem==188|sh11_pdsys$PDSystem==199)]<- 1 #Baxter product

sh11_pdsys$brand[which(sh11_pdsys$PDSystem==201|sh11_pdsys$PDSystem==202|sh11_pdsys$PDSystem==203|sh11_pdsys$PDSystem==204|sh11_pdsys$PDSystem==205|sh11_pdsys$PDSystem==206|sh11_pdsys$PDSystem==207|sh11_pdsys$PDSystem==288|sh11_pdsys$PDSystem==299)]<- 2 #Fresenius 

sh11_pdsys$brand[which(sh11_pdsys$PDSystem==301|sh11_pdsys$PDSystem==388|sh11_pdsys$PDSystem==399)]<- 3 #Lucenxia

sh11_pdsys$brand[which(is.na(sh11_pdsys$brand))]<-9
```

##### Fork to competing risk branch for peritonitis as outcome 
```{r wrangle_cmprsk_peri, eval=T, echo=F, cache=F, results=FALSE, warning=FALSE, message=FALSE}
# need to redo the steps in all_data EXCEPT dropping patients with PtOutcomeID==0
sh_final<-merge(sh7_8, sh_demo.fin, by="PatientID", all = TRUE)

sh_final<-arrange(sh_final, PatientID, PtDateOutcome)

## change column order 
##  | col_order = c("PatientID", "PtDateOutcome","PtOutcomeID", "event", "PtDateCommence", "PtDateCentre", "ESRFDateFirstRx", "day_diag", "day_pd") 

sh_final<-sh_final %>%
  relocate(PtDateCommence, .after=PatientID) %>%
  relocate(ESRFDateFirstRx, .after=PtDateCommence) %>%
  relocate(PtDateOutcome, .after = ESRFDateFirstRx) %>%
  relocate(PtOutcomeID, .after = PtDateOutcome) %>%
  relocate(day_diag, .after = PtOutcomeID) %>%
  relocate(day_pd, .after=day_diag) %>%
  relocate(event, .after=day_pd)

#If we don't drop PtOutcomeID==0, many rows will remain as NA

sh_final$PtDateOutcome[which(is.na(sh_final$PtDateOutcome))]<- as.Date("2018-12-31")
sh_final$event[which(is.na(sh_final$event))]<- 0 
sh_final$PtOutcomeID[which(is.na(sh_final$PtOutcomeID))]<- 0

# create t.start, tstop, time 
sh_final<-sh_final %>%
  mutate(t.start2=0) %>%
  relocate(t.start2, .after=event) %>%
  mutate(t.stop=PtDateOutcome - PtDateCommence) %>%
  relocate(t.stop, .after=t.start2)

sh_final<-sh_final %>%
  group_by(PatientID) %>%
  mutate(Index=row_number()) %>%
  relocate(Index, .after=t.stop) 

sh_final<-sh_final %>%
  group_by(PatientID) %>%
  mutate(t.start = ifelse(Index>1, lag(t.stop), 0)) %>%
  relocate(t.start, .before=t.stop)
  
 ## sh_final$t.start[sh_final$Index>1]<-lead(sh_final$t.stop) #this doesn't work 

## create a third duration variable 'time' 
sh_final['time']<-sh_final$t.stop - sh_final$t.start
sh_final<-relocate(sh_final, time, .after = t.stop)



sh_final <- sh_final %>%
  relocate(PtSex, .after=Index) %>%
  relocate(PtRace, .after=PtSex) %>%
  relocate(Age, .after=PtRace)

sh_final <- sh_final %>%
  mutate(death=ifelse(PtOutcomeID==1, 1, 0)) %>%
  relocate(death, .after=Age)

# check the category of Patient Outcome 
#plot_bar(sh_final$PtOutcomeID)  #shows 7 categories
                                #12: peritonitis, 1: death, 13: ESI, 
                                #2: HD, 8: recover kidney function,                                 
                                #6: lost to f/up, 3: transplanted 

comp <- sh_final #prepare a new datasets for wrangling for cox

#create a new variable "status" 
comp['status']<- NA
comp <- comp %>% 
  relocate(status, .after= PtOutcomeID)

comp$status[which(comp$PtOutcomeID==0)]  <-0
comp$status[which(comp$PtOutcomeID==12)] <-1
comp$status[which(comp$PtOutcomeID==2)]  <-2 #death & HD are denoted competing risks
comp$status[which(is.na(comp$status2))]<-0 
comp$status<-as.factor(comp$status2)


comp <- comp %>% 
  arrange(PatientID, PtDateOutcome) %>%
  mutate(Index=row_number()) %>%
  relocate(Index, .after=status) 

#find out which subject has index==1 and outcome==0 -> need to remove specifically these rows
comp2<-comp %>% 
  filter(n()!=1)

which(comp2$Index==1 & comp2$PtOutcomeID ==0)   #no subject had outcome==0 as first outcome event 
                                                ## comp2 has no more value 

comp <- comp %>% 
  arrange(PatientID, PtDateOutcome)  %>%
  filter(Index==1)

##  | drop PtHeight and PtWeight (for now since too many missing data)
comp <- subset(comp, select = -c(t.start2, NotifID, PtRRTID, SDPID, SDPID.x, PtARID, NotifID.x, PtRRTID.x, givendate.x,NotifID.y, PtRRTID.y, PtRenalSpecify, PtRenal2, PtRenalSpecify2, givendate.y, PtHeight, PtWeight, Modality))


######
######
### fill in PtAssist arbitrily then verify later 
######
######
comp$PtAssist[which(is.na(comp$PtAssist))]<-1
comp$PtAssist[which(comp$PtAssist==9999)]<-1


comp[is.na(comp)]<-0 #this step will cause many missing data in PtAssist to become "0"

#convert day_diag and day_pd to months 
comp['month_diag']<-comp$day_diag/30
comp['month_pd']<- comp$day_pd/30

######## Convert SDPID.y to Site (arranged by State) ######## 
comp['site']<- NA
comp$site[which(comp$SDPID.y==330)]<-2
comp$site[which(comp$SDPID.y==430)]<-3
comp$site[which(comp$SDPID.y==530 | comp$SDPID.y==630)]<-4
comp$site[which(comp$SDPID.y==830 | comp$SDPID.y==831)]<-5
comp$site[which(comp$SDPID.y==730 | comp$SDPID.y==4330 | comp$SDPID.y==108130)]<-6
comp$site[which(comp$SDPID.y==930)]<-7
comp$site[which(comp$SDPID.y==1030 | comp$SDPID.y==2330 | comp$SDPID.y==3030 | comp$SDPID.y==7731)]<-8
comp$site[which(comp$SDPID.y==1330)]<-9
comp$site[which(comp$SDPID.y==1230 | comp$SDPID.y==3130 | comp$SDPID.y==3730 | comp$SDPID.y==5030 | comp$SDPID.y==7130 | comp$SDPID.y==143730)]<-10
comp$site[which(comp$SDPID.y==1130 | comp$SDPID.y==8530 | comp$SDPID.y==106530)]<-11
comp$site[which(comp$SDPID.y==1430 | comp$SDPID.y==1530 | comp$SDPID.y==1630)]<-12
comp$site[which(comp$SDPID.y==1730 | comp$SDPID.y==1930)]<-13
comp$site[which(comp$SDPID.y==130)]<-14

comp$site <- as.factor(comp$site)

## combine the dataset with sh11_pdsys 
comp <- merge(comp, sh11_pdsys, by="PatientID", all=TRUE)


# create a new variable "dm" to denote primary renal disease of diabetes 
comp <- comp %>%
  mutate(dm=ifelse(PtRenal1==2, 1,0)) %>%
  relocate(dm, .after=PtAssist)

# create a new variable "Age2" to make it similar to Cox PH model 
library(AMR) 
comp <- comp %>%
  mutate(Age2=Age)

comp$Age2 <- age_groups(comp$Age, split_at=c(18,30,40,50,60,70,80,90))

comp$Age2<-factor(comp$Age2, ordered = FALSE)
  


# create a new variable "m.time" to denote time in month 
comp['m.time']<- comp$time/30

comp = comp %>%
  relocate(m.time, .after=time)

## double check the data structure of the "comp" data frame 
str(comp)   #status     : numeric 
            #day_diag   : difftim 
            #day_pd     : difftime 
            #month_diag : difftime
            #month_pd   : difftime
            #event      : numeric 
            #time       : difftime
            #death      : numeric 
            #PtAssist   : numeric 
            #dm         : numeric 
            #comob.hpt  : numeric 
            #comob.dm   : numeric 
            #brand      : numeric 
            #Age_group  : numeric

# Change variables back to factor 
a <- c("status", "event", "death", "PtAssist", "dm", "comob.hpt", "comob.dm","brand", "Age2")
comp[,a]<-lapply(comp[,a], factor)

# change variable from difftime to numeric 
b <- c("day_diag", "day_pd", "month_diag", "month_pd", "time", "m.time")
comp[,b]<-sapply(comp[,b], as.numeric)


```

```{r ff_final, eval=T, echo=F, cache=F, results=F, warning=F, message=F}


# Recode the variable 
library(finalfit)
library(dplyr)
library(forcats) #solve problems with factor variables 
library(kableExtra)

#for the vignette, pre-specify table output (can be modified for poster)
mykable = function(x){
  knitr::kable(x, row.names = FALSE, align = c("l","l","r","r","r","r","r","r","r")) %>%
    kable_styling(font_size=35)
}


#data use here is "comp" 
  #status == 0, no outcome 
  #status == 1, peritonitis 
  #status == 2, death or HD as competing risk 
  #variables included: PtRace, Age_group, PtAssist, dm, brand, site 
comp = comp %>% 
  mutate(
    #overall survival 
    status_os = ifelse(status == 1, 1, #peritonitis
                                      0), #no peritonitis 
    
    #disease specific survival 
    status_dss = case_when(
      status == 0 ~ 0, #no event  
      status == 1 ~ 1, #peritonitis 
      TRUE ~ 0),    #Other causes are censored 
    
    #competing risk regression 
    status_crr = case_when(
      status == 0 ~ 0, #no event 
      status == 1 ~ 1, #peritonitis 
      status == 2 ~ 2,
      TRUE ~ 0), #competing risk 
    
    #Label and recode other variable 
    Age2 = factor(Age2) %>%
      fct_recode("18-29" = "18-29", 
                 "30-39" = "30-39",
                 "40-49" = "40-49",
                 "50-59" = "50-59",
                 "60-69" = "60-69") %>%
      ff_label("Age Group"),
    PtRace = factor(PtRace) %>%
      fct_recode("Malay" = "1", 
                 "Chinese" = "2",
                 "Indian" = "3", 
                 "Bumi Sabah" = "5", 
                 "Bumi Sarawak" = "6", 
                 "Other Msian" = "19") %>%
      ff_label("Race"), 
    PtAssist = factor(PtAssist) %>%
      fct_recode("No help" = "1", 
                 "Partial" = "2", 
                 "Dependent" = "3") %>%
      ff_label("CAPD Assistance"), 
    dm = factor(dm) %>%
      fct_recode("Diabetic" = "1", 
                 "Non-diabetic" = "0") %>%
      ff_label("Primary Renal Disease"), 
    brand = factor(brand) %>%
      fct_recode("Baxter" = "1", 
                 "FMC" = "2", 
                 "Lucenxia" = "3") %>%
      ff_label("PD brand"), 
    site = factor(site) %>%
      fct_recode("Kedah" = "2", 
                 "Penang" = "3",
                 "Perak" = "4", 
                 "N.Sembilan" = "5", 
                 "Selangor" = "6", 
                 "Melaka" = "7", 
                 "Johor" = "8", 
                 "Kelantan" = "9", 
                 "Terengganu" = "10", 
                 "Pahang" = "11", 
                 "Sabah" = "12", 
                 "Sarawak" = "13", 
                 "KL" = "14") %>%
      ff_label("Treatment Centre")
    )

#relocate the status columns 
comp = comp %>%
  relocate(status_os, .after=status) %>%
  relocate(status_dss, .after=status_os) %>%
  relocate(status_crr, .after=status_dss)

#KAPLAN MEIER SURVIVAL ESTIMATORS 
library(survival)

survival_object = comp %$% #beware of the exposition paramter, and NOT %>%
  Surv(time = m.time, status_os)

#explore 
head(survival_object) # + marks censoring, in this case "no peritonitis" 

# Expressing time in months 
survival_object = comp %$%
  Surv(time/30, status_os)

# KM ANALYSIS FOR THE WHOLE COHORT 

# to plot survival stratified by  single grouping variable, substitute "survival_object~1" by "survival_object~factor" 

#overall survival in whole cohort 
 my_survfit = survfit(survival_object ~ 1, data = comp)
 my_survfit  #1149 patients, 418 events 

#Life table 
# A tabular form of KM plot that shows survival as a proportion, together with confidence limits 

summary(my_survfit, times = c(0, 12, 24, 36, 48, 60))


# Kaplan Meier Plot 
dependent_os = "Surv(time/30, status_os)"
explanatory = c("brand")

# comp %>%
#   surv_plot(dependent_os, explanatory, pval=TRUE)


# COX PROPORTIONAL HAZARD REGRESSION 

#Univariable and multivariable models 
dependent_os = "Surv(m.time, status_os)"
dependent_dss = "Surv(m.time, status_dss)"
dependent_crr = "Surv(m.time, status_crr)"
explanatory = c("PtRace", "Age2", "PtAssist", "dm", "brand", "site")

comp %>% 
  finalfit(dependent_os, explanatory) # %>%
  # mykable() #for vignette only 

#the labelling of the final table can be easily adjusted as desired 
comp %>%
  finalfit(dependent_os, explanatory, add_dependent_label = FALSE) %>%
  rename("Overall survival" = label) %>%
  rename(" " = levels) %>%
  rename("  " = all) %>%
  mykable()

#Reduced model 
#if you are using a backwards selection approach or similar, a reduced model can be directly specified and compared. The full model can be kept or dropped 

explanatory_multi = c("PtRace", "dm", "brand", "site")

comp %>%
  finalfit(dependent_os, explanatory, explanatory_multi, keep_models = TRUE) %>%
  mykable()

#Testing for porportional hazard
#the cox.zph() function from the survival package allows us to test this assumption for each variable. The plot of scaled Schoenfeld residuals should be a horizontl line. The included hypothesis test identifies whether the gradient differs from zero for each variables. No variable significantly differs from zero at the 5% significance level 

# explanatory = c("PtRace", "Age2", "PtAssist", "dm", "brand", "site")  #here Age got issue, might need to    change to Age2 - later
# comp %>% 
#   coxphmulti(dependent_os, explanatory) %>%
#   cox.zph() %>%
#   {zph_result <<- .} %>%
#   plot(var=5)

# zph_result


#Stratified models
# including a strata() term will result in a separate baseline hazard function being fit for each level in the stratification variable. It will no longer be possible to make direct inference on the effect associated with that variable 

explanatory = c("PtRace", "Age2", "PtAssist", "dm", "brand", "site", "strata(DataYear)") #DataYear might not be the the best strata() term, but just to illustrate

comp %>% 
  finalfit(dependent_os, explanatory) %>%
  mykable()


#Correlated Groups of Observations 
# to account for any higher structure in the data, e.g. patients are clustered within particular hospitals 
# Two broad approach to dealing with correlated groups of observatio n

# cluster() term: implies a generalised estimating equation (GEE) approach. A standard Cox PH model is fitted but the standard errors of the estimated hazard ratios are adjusted to account for correlations 

# frailty() term: implies a mixed effect model, where specific random effects term(s) are directly incorporated into the model 

# Both approaches achieve the same goal in different ways. the latter approach is favoured because of its flexibility and own preference. Note: cluster() and frailty() cannot be combined in the same model 

#Cluster model 
explanatory = c("PtRace", "Age2", "PtAssist", "brand", "dm", "cluster(site)")
comp %>% 
  finalfit(dependent_os, explanatory) %>%
  mykable()


#frailty model 
explanatory = c("PtRace", "Age2", "PtAssist", "brand", "dm", "frailty(site)")
comp %>% 
  finalfit(dependent_os, explanatory) %>%
  mykable()


#Hazard ratio plot 
#  comp %>% 
#    hr_plot(dependent_os, explanatory)

 
#COMPETING RISK REGRESSION 

#the Fine & Gray model is one option to deal with competing risk. The crr() syntax differs from survival::coxph() but finalfit brings these together 

#it uses the finalfit::ff_merge() function, which can join any number of models together 
 
explanatory = c("PtRace", "Age2", "PtAssist", "brand", "dm", "site")
dependent_dss = "Surv(m.time, status_dss)"
dependent_crr = "Surv(m.time, status_crr)"

  comp %>% 
    #summary table 
  summary_factorlist(dependent_dss, explanatory, column = TRUE, fit_id = TRUE) %>% 
  
    #Cox PH univariable 
      ff_merge(
        comp %>% 
          coxphuni(dependent_dss, explanatory) %>%
          fit2df(estimate_suffix = " (DSS CPH univariable)")
      ) %>%
    #Cox PH multivariable 
      ff_merge(
        comp %>% 
          coxphmulti(dependent_dss, explanatory) %>%
          fit2df(estimate_suffix = " (DSS CPH multivariable)")
      ) %>%
  
    #Fine & Gray Competing Risk model 
      ff_merge(
        comp %>% 
          crrmulti(dependent_crr, explanatory) %>% 
          fit2df(estimate_suffix = " (competing risks multivariable")
      ) %>% 
  
  select(-fit_id, -index) %>% 
  dependent_label(comp, "Survival") %>%
    mykable()


### Only CPH and competing risk   
explanatory = c("Age2","PtAssist", "brand", "site")
dependent_dss = "Surv(m.time, status_dss)"
dependent_crr = "Surv(m.time, status_crr)"

tab1 <- comp %>% 
    #summary table 
  summary_factorlist(dependent_dss, explanatory, column = TRUE, fit_id = TRUE) %>% 
  
    #Cox PH multivariable 
      ff_merge(
        comp %>% 
          coxphmulti(dependent_dss, explanatory) %>%
          fit2df(estimate_suffix = " (DSS CPH multivariable)")
      ) %>%
  
    #Fine & Gray Competing Risk model 
      ff_merge(
        comp %>% 
          crrmulti(dependent_crr, explanatory) %>% 
          fit2df(estimate_suffix = " (competing risks multivariable")
      ) %>% 
  
  select(-fit_id, -index) %>% 
  dependent_label(comp, "Survival") %>%
    mykable()


```


# Abstract 

# Introduction 








Literature review 

1. Chen et al 2019 

- clarify the risk factors of technique failure and outcomes (including PD-related peritonitis and mortality)
- multivariate analysis model (with subdistribution hazard)
- DM was a risk factor for peritonitis
- Female & higher serum albumin were associated with lower risks of technique failure 
- Gram-negative & polymicrobial infection increased the technique failure rate
- Female associated with overall mortality 
- Higher weekly urea clearance (Kt/V) and weekly creatinine clearance (WCCr) associated with a lower risk of mortality 

2. Evans et al 2010 
- Must cite paper 

3. Kim et al 2017 
- To evaluate the influence of education on the development of peritonitis, technique failure, and overall mortality 
- Cox PH model with adjustment (enter method) used to examine the independent risk factors related to all-cause mortality and time to first peritonitis. 
- Fine and Gray's model used to examine the difference in technique failure rates (between education levels?)

4. Ong et al 2016 
- Risk factors of PD-related in Malaysia multi-racial Asian population 
- Risk factors included (increase risk): severe obesity, hypoalbuminemia, *Staph. aureus* nasal carriage, FMC system, 
- Risk faactors (lower risk) automated PD vs standard PD, centre with patient-staff ratio 15-29.9 or >=30, prevalent patients, exit site care with antibiotics 

# Method

*Study design* 

*Setting* 

*Participants* 

*Variables* 

*Data source* 

*Bias* 

*Study size* 

*Quantitative variables* 

*Kaplan-Meier survival analysis* 
The standard Kaplan-Meier approach was used. The survival probability by time *t* referred to the risk of failure from peritonitis while all other causes of failure were absent. The cumulative incidence of peritonitis was calculated by the formula $CI=1-KM$

*Competing risk survival analysis* 
The event of interest was peritonitis, and competing risk events were death, transfer to HD, renal transplant, and partial recovery of kidney function prior to the first peritonitis. The CI functions for individual competing risk event were calculated using **????** approach. 

*Regression* 
Data from 1149 patients were included in the competing risk regression analysis. The aim was to identify risk factors (????)


*Software* 
The open-source R (version 3.6.0) was used. We used the packages *cmprsk* (version 2.2.9), *survival* (version 2.44.1.1), and *survminer* (version 0.4.5) for our analysis. 


# Results 

*Patient characteristcs* 
Patient characteristics are summarised in Table 1. 













# Analysis code


```{r cleaning,eval=T,echo=F, cache=T, results=F, warning=F, message=F}
library(tidyverse)
# | Phase 1 competing risk 
sh1_patient <- read_csv("Sh1_Patient.csv")
sh3_AR <- read_csv("Sh2_PtAnnualAR.csv")
sh4_notif <- read_csv("Sh4_PtNotif.csv")
sh5_comorb <- read_csv("Sh5_PtNotifComorb.csv")
sh6_lab <- read_csv("Sh6_PtNotifLab.csv")
sh7_outcome <- read_csv("Sh7_PtNotifOutcome.csv")
sh8_peri <- read_csv("Sh8_PtNotifPDPeri.csv")

sh9_periAB <- read_csv("Sh9_PtNotifPDPeriAB.csv")
sh10_pdperiperf <- read_csv("Sh10_PtNotifPDPeriPerf.csv")
sh11_pdsys <- read_csv("Sh11_PtNotifPDSystem.csv")
sh12_sero <- read_csv("Sh12_PtNotifSerology.csv")
sh13_sdp <- read_csv("Sh13_SDP.csv")

```

```{r sh1,eval=T,echo=T, cache=T, results=T, warning=F, message=F}
#Step 1: clarify the data structure 
str(sh1_patient) #Date format ok, others should be factor 

x<-c("PatientID","PtSex","PtCitizenship","PtRace","PtABO") 
sh1_patient[,x]<-lapply(sh1_patient[,x],factor) #use lapply() to iteratively convert to factor variables 

#Step 2: identify duplicate patients 
which(duplicated(sh1_patient$PatientID)) #all unique patients 

#Step 3: identify ANY missing data 
which(is.na(sh1_patient)) #no missing data

#Step 4: Explore & enrich the data 

### 4.1: histogram for Race, and Blood-group 
library(DataExplorer)
plot_bar(sh1_patient$PtSex)         #Male=1, Female=2
plot_bar(sh1_patient$PtCitizenship) #Malaysian=1
plot_bar(sh1_patient$PtRace)        #Malay=1, Chi=2, Ind=3, BumiSabah=4
                                    #BumiSarawak=5, OrgAsliSemj=7, OtherMsian=19, 
                                    #Foreigner=20
plot_bar(sh1_patient$PtABO)         #A=1, B=2, AB=3, O=4, NA=8888, missing=9999
#note: some patients blood-group (PtABO) is "not available"; variable will not be used.

### 4.2: convert dob to age ###
sh1_patient['Age']<-NA #create a new variable called "Age"
sh1_patient['givendate']<-as.Date("31-12-2018", format="%d-%m-%Y")

library(lubridate) #use the lubridate package for dates 
sh1_patient$Age<- interval(start=sh1_patient$PtDateBirth, end=sh1_patient$givendate)/duration(num = 1, units = "years") 

ggplot(sh1_patient, aes(x=Age)) + 
  geom_histogram()                  #histogram shows negative age 

#   |to identify which patients had negative age 
subset(sh1_patient, sh1_patient$givendate < sh1_patient$PtDateBirth) #70695,71570
sh1_patient$PtDateBirth[which(sh1_patient$PatientID==70695)]<- as.Date("1929-07-16") #no need to specify "format", or else will turn into NA 
sh1_patient$PtDateBirth[which(sh1_patient$PatientID==71570)]<- as.Date("1929-09-04")

#   |re-run age calculation 
sh1_patient$Age<- interval(start=sh1_patient$PtDateBirth, end=sh1_patient$givendate)/duration(num = 1, units = "years") 

ggplot(sh1_patient, aes(x=Age)) + 
  geom_histogram()                  #skewness to the left 

```

```{r sh3,eval=T,echo=T, cache=T, results=T, warning=F, message=F}

#Step 1: clarify the data structure 
str(sh3_AR) #PtARID, PatientID, NotifID, SDPID, PtAssist should be factor 
x<-c("PtARID", "PatientID", "NotifID","PtRRTID", "SDPID")
sh3_AR[,x]<-lapply(sh3_AR[,x],factor)

#Step 2: identify duplicate patients 
which(duplicated(sh3_AR$PatientID)) #show multiple entries from same patients 

#Step 3: identify ANY missing data 
which(is.na(sh3_AR))                #no missing data 
 
#Step 4: Explore & enrich the data 

## 4.1: arrange row according to PatientID then DataYear 
sh3_AR<-arrange(sh3_AR,PatientID, DataYear) #some patients did not return the following year -> died? lost to f/up?

## 4.2: identify which PtAssist changed / deteriorates 
#plot_bar(sh3_AR$PtAssist)
which(sh3_AR$PtAssist==8888) #26 rows has NA for PtAssist 
sh3_AR$PtAssist[which(sh3_AR$PtAssist==8888)]<-NA
which(is.na(sh3_AR$PtAssist)) #all 26 rows converted to NA

## 4.2.1: PtAssist Changed from the year before 
sh3_AR<-sh3_AR %>%
  group_by(PatientID) %>%
  mutate(Asst.change=ifelse(PtAssist==lag(PtAssist), 0, 1) ) #lag() the previous row

sh3_AR$Asst.change[which(is.na(sh3_AR$Asst.change))]<-0

## 4.2.2: PtAssist deteriorated from the year before 
sh3_AR<-sh3_AR %>%
  group_by(PatientID) %>%
  mutate(deteriorate=ifelse(PtAssist>lag(PtAssist), 1, 0)) #lag() the previous row

sh3_AR$deteriorate[which(is.na(sh3_AR$deteriorate))]<-0


## Create a new sub-data that contains latest DataYear
## Will sacrifice some details in PtAssist over time 
sh3_AR<-arrange(sh3_AR, PatientID, desc(DataYear))

library(data.table)
setDT(sh3_AR)
sh3_AR2<-sh3_AR[rowid(rleid(PatientID)) == 1]
sh3_AR2<-subset(sh3_AR2, select =  -c(8:9))

``` 

```{r sh4,eval=T,echo=T, cache=T, results=T, warning=F, message=F}

#Step 1: clarify the data structure 
str(sh4_notif)    #A few issues: 
                  #PatientID, NotifID, PtRRTID, SDPID - factor
                  #PtDateCommence, PtDateCentre, ESRFDateFirstRx - date 
                  #PtAssis - no.obs less than sh3_AR; ignore PtAssist in sh4_notif 
                  #Modality in 'character' but all should be PD pt; ignore 

x<-c("PatientID", "NotifID","PtRRTID", "SDPID")
sh4_notif[,x]<-lapply(sh4_notif[,x],factor)

func<- function(x) as.Date(x,format = "%d-%m-%y") #create a function to convert to Date
sh4_notif[,6:8]<-data.frame(lapply(sh4_notif[,6:8], func)) #lapply() to convert to Date

sh4_notif<-subset(sh4_notif, select= -c(15))

sh4_notif$PtRenal1<-as.factor(sh4_notif$PtRenal1)

#Step 2: identify duplicate patients 
sh4_notif<-arrange(sh4_notif, PatientID, PtDateCentre) #arrange by PtDateCentre; 
                                                       #repeated entry due to patient 
                                                       #changing treatment site 
which(duplicated(sh4_notif$PatientID))                 #69 repeat entries by PatientID 

## 2.1: subset patients who changed site -> ESRFDate not complete/missing 
sh4_subset<- sh4_notif %>%
  group_by(PatientID) %>%
  filter(n()!=1)               #this method produce 129 observations 
  
  
sh4_subset<-subset(sh4_notif, (duplicated(sh4_notif$PatientID) | duplicated(sh4_notif$PatientID, fromLast = TRUE)))     # this also produce 129 obs

#only one subject show different ESRFDateFirstRx; 66565 (SDPID 530)
sh4_notif$ESRFDateFirstRx[which(sh4_notif$PatientID==66565 & sh4_notif$SDPID==530)]<-as.Date("2013-03-20")

## Data exploration with sh4_subset 
##  | Aim to remove duplicate and see if any vital data gone missing 
sh4_subset$ESRFDateFirstRx[which(sh4_subset$PatientID==66565 & sh4_subset$SDPID==530)]<-as.Date("2013-03-20")
sh4_subset<-arrange(sh4_subset, PatientID, ESRFDateFirstRx)

##  | Use rleid() to generate run-length type group ID 
setDT(sh4_subset)
sh4_subset<-sh4_subset[rowid(rleid(PatientID)) == 1]

#75148, 75322, 75339, 75781, 75782: 5 pt missing PtRenal1
#double-check original dataset, confirm all 5 data fields missing

#before we make any changes to ESRFDateFirstRx, arrange() the dataset by PatientID, ESRFDateFirstRx
sh4_notif<-sh4_notif %>%
  group_by(PatientID) %>%
  arrange(sh4_notif, ESRFDateFirstRx, .by_group = TRUE)
#change ESRFDateFirstRx to character 
sh4_notif$ESRFDateFirstRx<-as.character(sh4_notif$ESRFDateFirstRx)

sh4_notif<- sh4_notif %>% 
  group_by(PatientID) %>%
  fill(ESRFDateFirstRx)
#change ESRFDateFirstRx back to date 
sh4_notif$ESRFDateFirstRx<-as.Date(sh4_notif$ESRFDateFirstRx)

#na_bool = is.na(sh4_notif$ESRFDateFirstRx)
#sh4_notif[na_bool,]$ESRFDateFirstRx=sh4_notif[c(FALSE, na_bool[-length(na_bool)]),]$ESRFDateFirstRx #shift the value up if ESRFDateFirstRx is NA

## 2.2: PtRenal1 
plot_bar(sh4_notif$PtRenal1) #quite some patient w/out valid reasons 

#Step 3: identify ANY missing data 
which(is.na(sh4_notif[,8])) #index 668 & 1217 have real missing data; 
                            #PatientID:72535,1003737
                            #should we assume that ESRF date starts with Dialysis date? Yes 

sh4_notif$ESRFDateFirstRx[which(sh4_notif$PatientID==72535)]<-as.Date("2014-10-12")
sh4_notif$ESRFDateFirstRx[which(sh4_notif$PatientID==1003737)]<-as.Date("2014-12-21")
                            

#Step 4: Explore & enrich the data 

## 4.1: create two new variables "times since diagnosis" and "time since 1st PD"
sh4_notif['day_diag'] <- NA
sh4_notif['day_pd'] <- NA 
sh4_notif['givendate']<-NA 

sh4_notif$givendate <- as.Date("31-12-2018", format="%d-%m-%Y")
sh4_notif$day_pd <- sh4_notif$givendate - sh4_notif$PtDateCommence
sh4_notif$day_diag <- sh4_notif$givendate - sh4_notif$ESRFDateFirstRx #the two cases above will have time calculated from PtDateCommence  
                                                      
## 4.2: remove all the duplicates
setDT(sh4_notif)
sh4_notif<-sh4_notif[rowid(rleid(PatientID)) == 1] #final dataset has 1149 obs

``` 

```{r sh5,eval=T,echo=T, cache=T, results=T, warning=F, message=F}

#Step 1: clarify the data structure 
str(sh5_comorb) #Issues identified 
                #PtNotifComorbidityID, PatientID, NotifID, PtRRTID, SDPID, Comorbidity - factor 
                #Drop DateDiagnosis, DateDiagnosis_Est, DateDiagnosis_Unk, DateResolve - all missing 

x<-c("PtNotifComorbidityID", "PatientID", "NotifID", "PtRRTID", "SDPID", "Comorbidity")
sh5_comorb[,x]<-lapply(sh5_comorb[,x], factor)

sh5_comorb<-subset(sh5_comorb, select = -c(8:11)) 

#Step 2: identify duplicate patients 
which(duplicated(sh5_comorb$PatientID)) #PatientID repeated for multiple comorbidities 
                                        #Should convert long to wide, each comorb is a variable 
                                        #See Step 4 for details 

#Step 3: identify ANY missing data 
which(is.na(sh5_comorb[,1:8])) #no missing data

#Step 4: Explore & enrich the data 

## 4.1: convert the column "Comorbidity" from long to wide
## | use the newer pivot_wider() from tidyverse/tidyr to make the conversion 
## | add a new column to indicate comorb is "present" 
sh5_comorb['comob_present']<-as.integer(1)
class(sh5_comorb$comob_present)

sh5_comorb <- subset(sh5_comorb, select = c(2,8,10))
sh5_comorb<-as_tibble(sh5_comorb)

sh5_comorb<-  sh5_comorb %>% 
  pivot_wider(
    names_from = Comorbidity,
    values_from = comob_present)

    
    #    values_fill = as.list("0")    #must force the "0" as the "list" type, or else it wouldnt work
  

## |Comorbidity type (change column name)
## comob.hpt, comob.dm, comob.peptic_ulcer, comob.other_cardiac, comob.cerebrovas, comob.chro_res, comob.ihd, comob.other, comob.cancer, comob.ccf, comob.impair_vision, comob.amputation, comob.chron_liver, comob.peri_vas, comob.psy, comob.GI, comob.cvd, comob.tb, comob.ment_retard, comob.dementia, comob.sub_abuse, 

## rename all columns 

names(sh5_comorb)<- c("PatientID", "comob.hpt", "comob.dm", "comob.peptic_ulcer", "comob.other_cardiac", "comob.cerebrovas", "comob.chro_res", "comob.ihd", "comob.other", "comob.cancer", "comob.ccf", "comob.impair_vision", "comob.amputation", "comob.chron_liver", "comob.peri_vas", "comob.psy", "comob.GI", "comob.cvd", "comob.tb", "comob.ment_retard", "comob.dementia", "comob.sub_abuse")

sh5_comorb$comob.dm[which(sh5_comorb$comob.dm=="NULL")]<-"0"
sh5_comorb$comob.hpt[which(sh5_comorb$comob.hpt=="NULL")]<-"0"
sh5_comorb$comob.peptic_ulcer[which(sh5_comorb$comob.peptic_ulcer=="NULL")]<-"0"
sh5_comorb$comob.other_cardiac[which(sh5_comorb$comob.other_cardiac=="NULL")]<-"0"
sh5_comorb$comob.cerebrovas[which(sh5_comorb$comob.cerebrovas=="NULL")]<-"0"
sh5_comorb$comob.chro_res[which(sh5_comorb$comob.chro_res=="NULL")]<-"0"
sh5_comorb$comob.ihd[which(sh5_comorb$comob.ihd=="NULL")]<-"0"
sh5_comorb$comob.other[which(sh5_comorb$comob.other=="NULL")]<-"0"
sh5_comorb$comob.cancer[which(sh5_comorb$comob.cancer=="NULL")]<-"0"
sh5_comorb$comob.ccf[which(sh5_comorb$comob.ccf=="NULL")]<-"0"
sh5_comorb$comob.impair_vision[which(sh5_comorb$comob.impair_vision=="NULL")]<-"0"
sh5_comorb$comob.amputation[which(sh5_comorb$comob.amputation=="NULL")]<-"0"
sh5_comorb$comob.chron_liver[which(sh5_comorb$comob.chron_liver=="NULL")]<-"0"
sh5_comorb$comob.peri_vas[which(sh5_comorb$comob.peri_vas=="NULL")]<-"0"
sh5_comorb$comob.psy[which(sh5_comorb$comob.psy=="NULL")]<-"0"
sh5_comorb$comob.GI[which(sh5_comorb$comob.GI=="NULL")]<-"0"
sh5_comorb$comob.cvd[which(sh5_comorb$comob.cvd=="NULL")]<-"0"
sh5_comorb$comob.tb[which(sh5_comorb$comob.tb=="NULL")]<-"0"
sh5_comorb$comob.ment_retard[which(sh5_comorb$comob.ment_retard=="NULL")]<-"0"
sh5_comorb$comob.dementia[which(sh5_comorb$comob.dementia=="NULL")]<-"0"
sh5_comorb$comob.sub_abuse[which(sh5_comorb$comob.sub_abuse=="NULL")]<-"0"

sh5_comorb$PatientID<-as.character(sh5_comorb$PatientID)

sh5_comorb$comob.dm[which(sh5_comorb$comob.dm=="1")]<-1
sh5_comorb$comob.hpt[which(sh5_comorb$comob.hpt=="1")]<-1
sh5_comorb$comob.hpt[which(sh5_comorb$PatientID=="78556")]<-1
sh5_comorb$comob.peptic_ulcer[which(sh5_comorb$comob.peptic_ulcer=="1")]<-1
sh5_comorb$comob.other_cardiac[which(sh5_comorb$comob.other_cardiac=="1")]<-1
sh5_comorb$comob.cerebrovas[which(sh5_comorb$comob.cerebrovas=="1")]<-1
sh5_comorb$comob.chro_res[which(sh5_comorb$comob.chro_res=="1")]<-1
sh5_comorb$comob.ihd[which(sh5_comorb$comob.ihd=="1")]<-1
sh5_comorb$comob.other[which(sh5_comorb$comob.other=="1")]<-1
sh5_comorb$comob.cancer[which(sh5_comorb$comob.cancer=="1")]<-1
sh5_comorb$comob.ccf[which(sh5_comorb$comob.ccf=="1")]<-1
sh5_comorb$comob.impair_vision[which(sh5_comorb$comob.impair_vision=="1")]<-1
sh5_comorb$comob.amputation[which(sh5_comorb$comob.amputation=="1")]<-1
sh5_comorb$comob.chron_liver[which(sh5_comorb$comob.chron_liver=="1")]<-1
sh5_comorb$comob.peri_vas[which(sh5_comorb$comob.peri_vas=="1")]<-1
sh5_comorb$comob.psy[which(sh5_comorb$comob.psy=="1")]<-1
sh5_comorb$comob.GI[which(sh5_comorb$comob.GI=="1")]<-1
sh5_comorb$comob.cvd[which(sh5_comorb$comob.cvd=="1")]<-1
sh5_comorb$comob.tb[which(sh5_comorb$comob.tb=="1")]<-1
sh5_comorb$comob.ment_retard[which(sh5_comorb$comob.ment_retard=="1")]<-1
sh5_comorb$comob.dementia[which(sh5_comorb$comob.dementia=="1")]<-1
sh5_comorb$comob.sub_abuse[which(sh5_comorb$comob.sub_abuse=="1")]<-1


sh5_comorb$comob.dm<-as.numeric(sh5_comorb$comob.dm)
sh5_comorb$comob.hpt<-as.numeric(sh5_comorb$comob.hpt) #Id 78556
sh5_comorb$comob.peptic_ulcer<-as.numeric(sh5_comorb$comob.peptic_ulcer)
sh5_comorb$comob.other_cardiac<-as.numeric(sh5_comorb$comob.other_cardiac)
sh5_comorb$comob.cerebrovas<-as.numeric(sh5_comorb$comob.cerebrovas)
sh5_comorb$comob.chro_res<-as.numeric(sh5_comorb$comob.chro_res)
sh5_comorb$comob.ihd<-as.numeric(sh5_comorb$comob.ihd)
sh5_comorb$comob.other<-as.numeric(sh5_comorb$comob.other)
sh5_comorb$comob.cancer<-as.numeric(sh5_comorb$comob.cancer)
sh5_comorb$comob.ccf<-as.numeric(sh5_comorb$comob.ccf)
sh5_comorb$comob.impair_vision<-as.numeric(sh5_comorb$comob.impair_vision)
sh5_comorb$comob.amputation<-as.numeric(sh5_comorb$comob.amputation)
sh5_comorb$comob.chron_liver<-as.numeric(sh5_comorb$comob.chron_liver)
sh5_comorb$comob.peri_vas<-as.numeric(sh5_comorb$comob.peri_vas)
sh5_comorb$comob.psy<-as.numeric(sh5_comorb$comob.psy)
sh5_comorb$comob.GI<-as.numeric(sh5_comorb$comob.GI)
sh5_comorb$comob.cvd<-as.numeric(sh5_comorb$comob.cvd)
sh5_comorb$comob.tb<-as.numeric(sh5_comorb$comob.tb)
sh5_comorb$comob.ment_retard<-as.numeric(sh5_comorb$comob.ment_retard)
sh5_comorb$comob.dementia<-as.numeric(sh5_comorb$comob.dementia)
sh5_comorb$comob.sub_abuse<-as.numeric(sh5_comorb$comob.sub_abuse)
sh5_comorb$PatientID<-as.numeric(sh5_comorb$PatientID)



``` 

```{r sh6,eval=T,echo=T, cache=T, results=T, warning=F, message=F}
## | Note: laboratory results had too many missing data; ignore this sheet at this moment 

#Step 1: clarify the data structure 

#Step 2: identify duplicate patients 

#Step 3: identify ANY missing data 

#Step 4: Explore & enrich the data 

``` 

```{r sh7,eval=T,echo=T, cache=T, results=T, warning=F, message=F}

#Step 1: clarify the data structure 
str(sh7_outcome)  #Similar issues as previous sheets 
                  #PtNotifOutcomeID, PatientID, NotifID, PtRRTID, SDPID - factor 
                  #Other outcome variables should be in "factor" format but leave them be numeric for now 

x<-c("PtNotifOutcomeID", "PatientID", "NotifID", "PtRRTID", "SDPID", "PtOutcomeID")
sh7_outcome[,x]<-lapply(sh7_outcome[,x], factor)

sh7_outcome$PtDateOutcome <- as.Date(sh7_outcome$PtDateOutcome, format = "%d-%m-%y")
sh7_outcome$PtDateOutcomeNotified <- as.Date(sh7_outcome$PtDateOutcomeNotified, format="%d-%m-%y")


#Step 2: identify duplicate patients 
sh7_outcome<-arrange(sh7_outcome, PatientID, PtDateOutcome)

which(duplicated(sh7_outcome$PatientID))  #identify 172 duplicated records - multiple outcomes by the same patient 

#Step 3: identify ANY missing data 

## 3.1: important variable is the "PtDateOutcome. Missing values in PtDateOutcomeNotified are ignored
which(is.na(sh7_outcome$PtDateOutcome))  #no missing data 


#Step 4: Explore & enrich the data 

## 4.1: check if "death" (terminal event) occurs befor any other outcome; death ==1 
sh7_outcome2<-subset(sh7_outcome, !sh7_outcome$PtOutcomeID==0)
sh7_outcome2<- sh7_outcome2[,1:9]

sh7_outcome3<- sh7_outcome2 %>%
  group_by(PatientID, desc(PtOutcomeID)) %>%
  filter(n()!=1) %>%
  mutate(wrong=ifelse((PtDateOutcome > lag(PtDateOutcome)),1,0)) #No event occurs after death; 

## 4.2: drop the column "modality" 
sh7_outcome<-subset(sh7_outcome, select =  -c(5))

## 4.3: drop the row with PtOutcomeID == 5 (move to other centre)
sh7_outcome<-subset(sh7_outcome, !(PtOutcomeID==5 | PtOutcomeID==0))
plot_bar(sh7_outcome$PtOutcomeID)

## 4.3: shorten the column 
sh7_outcome<-subset(sh7_outcome, select = c(1:9))
``` 

```{r sh8,eval=T,echo=T, cache=T, results=T, warning=F, message=F}

#Step 1: clarify the data structure 
str(sh8_peri)     #A few issues
                  #PtNotifPDPeriID, PatientID, NotifID, PtRRTID, SDPID, PtARID, PeritonitisType - Factor 
                  #DatePeri - Date

x<-c("PtNotifPDPeriID", "PatientID", "NotifID", "PtRRTID", "SDPID", "PtARID")
sh8_peri[,x]<-lapply(sh8_peri[,x], factor)

sh8_peri$DatePeri<-as.Date(sh8_peri$DatePeri, format="%d-%m-%y")

#Step 2: identify duplicate patients 
which(duplicated(sh8_peri$PatientID)) #541 repeated instances 

#Step 3: identify ANY missing data 
## | important variabls are DatePeri and PeritonitisType 
which(is.na(sh8_peri$DatePeri) | is.na(sh8_peri$PeritonitisType)) #no missing data

#Step 4: Explore & enrich the data 
## | arrange according to PatientID and DatePeri, and complete the remaining analysis in following sections

sh8_peri<-sh8_peri %>%
  arrange(PatientID, DatePeri)

sh8_peri<-subset(sh8_peri,select= c(1:9))


##  | some issues identified from preliminary analysis: 
##    1. Wrong PtDateOutcome (!) 70674, peri 2005-09-17;  75995 peri 2006-10-23; 
##    2. 70766, move to other centre, 2014-03-01; 
##    3. 72388, ESI 2014-09-01; 74135, peri 2014-05-19;76595 peri 2014-02-08 (date ESI is correct, need to check PtDateCommence)

# Step 1: Correct wrong data entry (PatientID == 70674, 75995)
sh8_peri$DatePeri[(sh8_peri$PatientID==70674 & sh8_peri$PtNotifPDPeriID==11619)]<-as.Date("2015-09-17")

sh8_peri$DatePeri[(sh8_peri$PatientID==75995 |sh8_peri$PtNotifPDPeriID==13695)]<-as.Date("2016-10-23") #this subject reported multipl ESI & peri on the same date 

# Step 2: Cannot verify date of peri, possibly experience before formally enrolled in dialysis program 
# to remove the incidents 
which(sh8_peri$PatientID==72388 & sh8_peri$PtNotifPDPeriID==10874) #index 612
sh8_peri<-sh8_peri[-612,]

which(sh8_peri$PatientID==74135 & sh8_peri$PtNotifPDPeriID==10309) #index 715
sh8_peri<-sh8_peri[-715,]

which(sh8_peri$PatientID==76595 & sh8_peri$PtNotifPDPeriID==11103) #index 1049
sh8_peri<-sh8_peri[-1049,]

# Step 3: Remove ESI from cases where ESI and Peritonitis co-exist on the SAME DATE 
sh8_peri2<- sh8_peri %>% 
  group_by(PatientID, DatePeri) %>%
  filter(n()!=1) %>%
  filter(DatePeri == lead(DatePeri))

x<-which(sh8_peri$DatePeri == lead(sh8_peri$DatePeri)) 

sh8_peri <- sh8_peri[-c(129,258,261,303,346,349,425,521,562,619,621,637,646,680,757,778,877,921,933,976,977,978,1031),]

# Double check if the removal is successful 
which(sh8_peri$DatePeri == lead(sh8_peri$DatePeri)) 

# Step 4: Many subjects have same date peritonitis with PtDateCommence 
#      4.1: Some subjects have same peritonitis date with death or transfer to HD 

#70408, 2014-05-12 +1
#70417, 2017-05-15 (same date as transfer to HD) -1
#70685, 2014-03-23  +1
#70687, 2014-04-27 +1
#70715, 2015-03-30 (same date as death) -1
#70854, 2017-06-17 (same date as death) -1
#71669, 2014-08-04 +1
#73942, 2014-08-04 +1
#74147, 2017-07-17 (same date as transfer to HD) -1
#76181, 2016-10-17 (same date as transfer to HD) -1
#76394, 2017-11-18 (same date as death) -1

sh8_peri$DatePeri[(sh8_peri$PatientID==70408 |sh8_peri$PtNotifPDPeriID==11016)]<-as.Date("2014-05-14")

sh8_peri$DatePeri[(sh8_peri$PatientID==70685 |sh8_peri$PtNotifPDPeriID==10254)]<-as.Date("2014-03-25")

sh8_peri$DatePeri[(sh8_peri$PatientID==70687 |sh8_peri$PtNotifPDPeriID==10260)]<-as.Date("2014-04-29")

sh8_peri$DatePeri[(sh8_peri$PatientID==71669 |sh8_peri$PtNotifPDPeriID==10705)]<-as.Date("2014-08-06")

sh8_peri$DatePeri[(sh8_peri$PatientID==73942 |sh8_peri$PtNotifPDPeriID==10250)]<-as.Date("2014-08-06")

sh8_peri$DatePeri[(sh8_peri$PatientID==70417 |sh8_peri$PtNotifPDPeriID==16027)]<-as.Date("2017-05-13")

sh8_peri$DatePeri[(sh8_peri$PatientID==70715 |sh8_peri$PtNotifPDPeriID==12029)]<-as.Date("2015-03-28")

sh8_peri$DatePeri[(sh8_peri$PatientID==70854 |sh8_peri$PtNotifPDPeriID==14790)]<-as.Date("2017-06-15")

sh8_peri$DatePeri[(sh8_peri$PatientID==74147 |sh8_peri$PtNotifPDPeriID==15678)]<-as.Date("2017-07-15")

sh8_peri$DatePeri[(sh8_peri$PatientID==76181 |sh8_peri$PtNotifPDPeriID==13873)]<-as.Date("2016-10-16")

sh8_peri$DatePeri[(sh8_peri$PatientID==76394 |sh8_peri$PtNotifPDPeriID==15339)]<-as.Date("2017-11-16")

## Step 5: Some patients have double entry 
#rep<-sh8_peri %>%
#  group_by(PatientID, PtDateOutcome) %>%
#  filter(n()!=1) %>% #a total of 5 unique subject id with 18 obs 
#  filter(PtDateOutcome==lead(PtDateOutcome)) #this filters down to 13 obs (correctly minus 5)




which(sh8_peri$PatientID==70685 & sh8_peri$PtNotifPDPeriID==10254)
which(sh8_peri$PatientID==71669 & sh8_peri$PtNotifPDPeriID==10705)
which(sh8_peri$PatientID==71669 & sh8_peri$PtNotifPDPeriID==13614)
which(sh8_peri$PatientID==71669 & sh8_peri$PtNotifPDPeriID==13615)
which(sh8_peri$PatientID==71669 & sh8_peri$PtNotifPDPeriID==13616)
which(sh8_peri$PatientID==71669 & sh8_peri$PtNotifPDPeriID==13617)
which(sh8_peri$PatientID==71669 & sh8_peri$PtNotifPDPeriID==13618)
which(sh8_peri$PatientID==71669 & sh8_peri$PtNotifPDPeriID==13619)
which(sh8_peri$PatientID==71669 & sh8_peri$PtNotifPDPeriID==13620)
which(sh8_peri$PatientID==73942 & sh8_peri$PtNotifPDPeriID==10250)
which(sh8_peri$PatientID==76181 & sh8_peri$PtNotifPDPeriID==12162)
which(sh8_peri$PatientID==76181 & sh8_peri$PtNotifPDPeriID==13872)
which(sh8_peri$PatientID==76394 & sh8_peri$PtNotifPDPeriID==15337)


#index 394,531,532,533,534,535,536,537,538,673,992,993,1016
sh8_peri <- sh8_peri[-c(394,531,532,533,534,535,536,537,538,673,992,993,1016),]
    
``` 

```{r sh7_8, eval=T, echo=F, cache=T, results=FALSE, warning=FALSE, message=FALSE} 

# combine both sh7 and sh8 into one outcome dataset 

## Step 1: change the PeritonitisType to PtOutcomeID, and DatePeri to PtDateOutcome 
##  | Peritonitis from 2-> 12; ESI from 1-> 13 
sh8_peri$PeritonitisType[sh8_peri$PeritonitisType==2]<- 12
sh8_peri$PeritonitisType[sh8_peri$PeritonitisType==1]<- 13

##  | Rename column 
sh8_peri<- rename(sh8_peri, PtOutcomeID=PeritonitisType)
sh8_peri<- rename(sh8_peri, PtDateOutcome=DatePeri)

##  | rbind both dataset with data.table()
library(data.table)
z = list(sh7_outcome, sh8_peri)
sh7_8<- rbindlist(z, use.names = TRUE, fill = TRUE)

#sh7_8<-merge(sh7_outcome, sh8_peri, by="PatientID", all = TRUE)

## Step 2: re-arrange the new combined dataset with PatientID 
sh7_8<-arrange(sh7_8, PatientID, PtDateOutcome)

which(is.na(sh7_8$PtOutcomeID) | is.na(sh7_8$PtDateOutcome)) #no missing data 

## Step 3: create event; t.start, t.stop, time cannot create without PtOutcomeDate 

sh7_8<-sh7_8 %>%
  mutate(event=ifelse(PtOutcomeID==12, 1,0)) %>%
  relocate(event, .after=PtOutcomeID)

sh7_8 <- subset(sh7_8,select =  c(2:8))

```

```{r sh_demo, eval=T, echo=F, cache=T, results=FALSE, warning=FALSE, message=FALSE}

# combine sh1, sh3, sh4, and sh5 to form a combined demographic dataset 

library(data.table)
#combine sh1_patient and sh5_comorb first 
sh_demo1<-merge(sh1_patient, sh5_comorb, by="PatientID", all = TRUE)
sh_demo2<-merge(sh_demo1, sh4_notif, by="PatientID", all=TRUE)
sh_demo.fin<-merge(sh3_AR2,sh_demo2, by="PatientID", all = TRUE)

sh_demo.fin<-arrange(sh_demo.fin, PatientID) #1149 obs 

```

##### stop here ! see which branch to fork 

```{r all_data, eval=T, echo=F, cache=T, results=FALSE, warning=FALSE, message=FALSE} 

# Combine sh_demo.fin and sh7_8 to make one final dataset 

# Step 1: understand the data sets 
## sh_demo.fin: 1149 obs, 51 variables 
## sh7_8:       2432 obs, 79 variables 

## use merge() to combine both datasets 
sh_final<-merge(sh7_8, sh_demo.fin, by="PatientID", all = TRUE)

sh_final<-arrange(sh_final, PatientID, PtDateOutcome)

## change column order 
##  | col_order = c("PatientID", "PtDateOutcome","PtOutcomeID", "event", "PtDateCommence", "PtDateCentre", "ESRFDateFirstRx", "day_diag", "day_pd") 

sh_final<-sh_final %>%
  relocate(PtDateCommence, .after=PatientID) %>%
  relocate(ESRFDateFirstRx, .after=PtDateCommence) %>%
  relocate(PtDateOutcome, .after = ESRFDateFirstRx) %>%
  relocate(PtOutcomeID, .after = PtDateOutcome) %>%
  relocate(day_diag, .after = PtOutcomeID) %>%
  relocate(day_pd, .after=day_diag) %>%
  relocate(event, .after=day_pd)

# drop rows when PtOutcomeID == 0 
sh_final<-subset(sh_final, !sh_final$PtOutcomeID==0)

# create t.start, tstop, time 
sh_final<-sh_final %>%
  mutate(t.start2=0) %>%
  relocate(t.start2, .after=event) %>%
  mutate(t.stop=PtDateOutcome - PtDateCommence) %>%
  relocate(t.stop, .after=t.start2)

sh_final<-sh_final %>%
  group_by(PatientID) %>%
  mutate(Index=row_number()) %>%
  relocate(Index, .after=t.stop) 

sh_final<-sh_final %>%
  group_by(PatientID) %>%
  mutate(t.start = ifelse(Index>1, lag(t.stop), 0)) %>%
  relocate(t.start, .before=t.stop)
  
 ## sh_final$t.start[sh_final$Index>1]<-lead(sh_final$t.stop) #this doesn't work 

## create a third duration variable 'time' 
sh_final['time']<-sh_final$t.stop - sh_final$t.start
sh_final<-relocate(sh_final, time, .after = t.stop)

```

##### Additional dataset from PD system and PD regimen 
```{r sh_11, eval=T, echo=F, cache=T, results=FALSE, warning=FALSE, message=FALSE}

sh11_pdsys <- read_csv("Sh11_PtNotifPDSystem.csv")


#Step 1: clarify the data structure 
str(sh11_pdsys) #PDRegimen and PDSystem were both numeric variables 
                #DateStart and DateEnd were character variables 

x<-c("PDRegimen", "PDSystem", "PatientID")
sh11_pdsys[,x]<-lapply(sh11_pdsys[,x], factor)

sh11_pdsys$DateStart <- as.Date(sh11_pdsys$DateStart, format = "%d-%m-%y")
sh11_pdsys$DateEnd <- as.Date(sh11_pdsys$DateEnd, format = "%d-%m-%y")

#set PDRegimen ==8888 or 9999 to NA 
sh11_pdsys$PDRegimen[which(sh11_pdsys$PDRegimen==8888 | sh11_pdsys$PDRegimen==9999)]<-NA

sh11_pdsys <- sh11_pdsys %>%
  group_by(PatientID) %>%
  arrange(PatientID, DateStart) %>%
  mutate(Index=row_number()) %>%
  relocate(Index, .after=PDRegimen)

which(is.na(sh11_pdsys$PDRegimen)&sh11_pdsys$Index==1) #1 (ID:1253), 79 (ID: 25299), 1601 (ID:70770), 1754 (ID: 71009)

#change PDRegimen into a factor variable 
sh11_pdsys$PDRegimen<- as.factor(sh11_pdsys$PDRegimen)

sh11_pdsys$PDRegimen[which(sh11_pdsys$PatientID==1253&sh11_pdsys$Index==1)]<-3
sh11_pdsys$PDRegimen[which(sh11_pdsys$PatientID==70770&sh11_pdsys$Index==1)]<-1
sh11_pdsys$PDRegimen[which(sh11_pdsys$PatientID==71009&sh11_pdsys$Index==1)]<-1
sh11_pdsys$PDRegimen[which(sh11_pdsys$PatientID==25299&sh11_pdsys$Index==1)]<-1

#set PDSystem ==8888 or 9999 to NA 
sh11_pdsys$PDSystem[which(sh11_pdsys$PDSystem==8888 | sh11_pdsys$PDSystem==9999)]<-NA

sh11_pdsys<- relocate(sh11_pdsys, PDSystem, .after = Index)

which(is.na(sh11_pdsys$PDSystem)&sh11_pdsys$Index==1) #1 (ID:1253), 79 (ID: 25299), 1601 (ID:70770), 1754 (ID: 71009)

sh11_pdsys$PDSystem[which(sh11_pdsys$PatientID==1253&sh11_pdsys$Index==1)]<-101
sh11_pdsys$PDSystem[which(sh11_pdsys$PatientID==70770&sh11_pdsys$Index==1)]<-101
sh11_pdsys$PDSystem[which(sh11_pdsys$PatientID==71009&sh11_pdsys$Index==1)]<-101
sh11_pdsys$PDSystem[which(sh11_pdsys$PatientID==25299&sh11_pdsys$Index==1)]<-101

## Keep only the first row 
sh11_pdsys <- sh11_pdsys %>%
  arrange(PatientID, Index) %>%
  filter(Index==1)

which(is.na(sh11_pdsys$PDRegimen) | is.na(sh11_pdsys$PDSystem)) #no missing data 

## trim the dataset to contain only PatientID, PDSystem, and PDRegimen 
sh11_pdsys<- select(sh11_pdsys, c("PatientID", "PDSystem", "PDRegimen"))

## create new variable for PDSystem as "brand" 
sh11_pdsys['brand']<-NA
sh11_pdsys$brand[which(sh11_pdsys$PDSystem==101|sh11_pdsys$PDSystem==102|sh11_pdsys$PDSystem==103|sh11_pdsys$PDSystem==104|sh11_pdsys$PDSystem==105|sh11_pdsys$PDSystem==106|sh11_pdsys$PDSystem==188|sh11_pdsys$PDSystem==199)]<- 1 #Baxter product

sh11_pdsys$brand[which(sh11_pdsys$PDSystem==201|sh11_pdsys$PDSystem==202|sh11_pdsys$PDSystem==203|sh11_pdsys$PDSystem==204|sh11_pdsys$PDSystem==205|sh11_pdsys$PDSystem==206|sh11_pdsys$PDSystem==207|sh11_pdsys$PDSystem==288|sh11_pdsys$PDSystem==299)]<- 2 #Fresenius 

sh11_pdsys$brand[which(sh11_pdsys$PDSystem==301|sh11_pdsys$PDSystem==388|sh11_pdsys$PDSystem==399)]<- 3 #Lucenxia

sh11_pdsys$brand[which(is.na(sh11_pdsys$brand))]<-9
```

##### Fork to KM & Cox branch 

```{r wrangle_cox, eval=T, echo=F, cache=T, results=FALSE, warning=FALSE, message=FALSE} 
# need to redo the steps in all_data EXCEPT dropping patients with PtOutcomeID==0
sh_final<-merge(sh7_8, sh_demo.fin, by="PatientID", all = TRUE)

sh_final<-arrange(sh_final, PatientID, PtDateOutcome)

## change column order 
##  | col_order = c("PatientID", "PtDateOutcome","PtOutcomeID", "event", "PtDateCommence", "PtDateCentre", "ESRFDateFirstRx", "day_diag", "day_pd") 

sh_final<-sh_final %>%
  relocate(PtDateCommence, .after=PatientID) %>%
  relocate(ESRFDateFirstRx, .after=PtDateCommence) %>%
  relocate(PtDateOutcome, .after = ESRFDateFirstRx) %>%
  relocate(PtOutcomeID, .after = PtDateOutcome) %>%
  relocate(day_diag, .after = PtOutcomeID) %>%
  relocate(day_pd, .after=day_diag) %>%
  relocate(event, .after=day_pd)

#If we don't drop PtOutcomeID==0, many rows will remain as NA

sh_final$PtDateOutcome[which(is.na(sh_final$PtDateOutcome))]<- as.Date("2018-12-31")
sh_final$event[which(is.na(sh_final$event))]<- 0 
sh_final$PtOutcomeID[which(is.na(sh_final$PtOutcomeID))]<- 0

# create t.start, tstop, time 
sh_final<-sh_final %>%
  mutate(t.start2=0) %>%
  relocate(t.start2, .after=event) %>%
  mutate(t.stop=PtDateOutcome - PtDateCommence) %>%
  relocate(t.stop, .after=t.start2)

sh_final<-sh_final %>%
  group_by(PatientID) %>%
  mutate(Index=row_number()) %>%
  relocate(Index, .after=t.stop) 

sh_final<-sh_final %>%
  group_by(PatientID) %>%
  mutate(t.start = ifelse(Index>1, lag(t.stop), 0)) %>%
  relocate(t.start, .before=t.stop)
  
 ## sh_final$t.start[sh_final$Index>1]<-lead(sh_final$t.stop) #this doesn't work 

## create a third duration variable 'time' 
sh_final['time']<-sh_final$t.stop - sh_final$t.start
sh_final<-relocate(sh_final, time, .after = t.stop)



sh_final <- sh_final %>%
  relocate(PtSex, .after=Index) %>%
  relocate(PtRace, .after=PtSex) %>%
  relocate(Age, .after=PtRace)

sh_final <- sh_final %>%
  mutate(death=ifelse(PtOutcomeID==1, 1, 0)) %>%
  relocate(death, .after=Age)

# check the category of Patient Outcome 
plot_bar(sh_final$PtOutcomeID)  #shows 7 categories
                                #12: peritonitis, 1: death, 13: ESI, 
                                #2: HD, 8: recover kidney function,                                 
                                #6: lost to f/up, 3: transplanted 

sh_cox <- sh_final #prepare a new datasets for wrangling for cox

# For patients who experience other outcomes before death, the outcomes should be remove (leaving death as the only outcome)
## Plan: rearrange row by PtDateOutcome and remove the remaining rows
## Rationale: if other outcomes (HD, death, transplanted, lost to f/up, recover kidney function) occurs before peritonitis, these will "hide" the subsequent risks - competing risk analysis
# For patients who had no competing outcomes, death will be terminal outcome, others will be censored

sh_cox <- sh_cox %>% 
  arrange(PatientID, PtDateOutcome)  %>%
  filter(Index==1)

## | use the newer pivot_wider() from tidyverse/tidyr to make the conversion 
## | add a new column to indicate primary renal disease is "present" 
sh_cox['prim.dis']<-as.integer(1)
class(sh_cox$prim.dis)

sh_cox <- sh_cox %>%
  pivot_wider(
    names_from = PtRenal1, 
    values_from = prim.dis
  )

##  | rename the new columns 
names(sh_cox)[63:73] <- c("toxic.nephropthy", "hpt","glomerulonephritis", "dm", "unknown", "obs.uropathy", "nil", "miss","others","hereditory.nephritis", "ADPKD")

##  | drop column nil and miss 
sh_cox<- sh_cox[,-(69:71)]

##  | drop PtHeight and PtWeight (for now since too many missing data)
sh_cox <- subset(sh_cox, select = -c(t.start2, NotifID, PtRRTID, SDPID, SDPID.x, PtARID, NotifID.x, PtRRTID.x, givendate.x,NotifID.y, PtRRTID.y, PtRenalSpecify, PtRenal2, PtRenalSpecify2, givendate.y, PtHeight, PtWeight, Modality))

sh_cox[is.na(sh_cox)]<-0

```

```{r Kaplan-Meier, eval=T, echo=F, cache=T, results=FALSE, warning=FALSE, message=FALSE}

library(survival)
library(survminer)

######################################
### Customise label for ggsurvplot ###
######################################

#Helper function to customize plot labels:

customize_labels <- function (p, font.title = NULL,
                              font.subtitle = NULL, font.caption = NULL,
                              font.x = NULL, font.y = NULL, font.xtickslab = NULL, font.ytickslab = NULL)
{
  original.p <- p
  if(is.ggplot(original.p)) list.plots <- list(original.p)
  else if(is.list(original.p)) list.plots <- original.p
  else stop("Can't handle an object of class ", class (original.p))
  .set_font <- function(font){
    font <- ggpubr:::.parse_font(font)
    ggtext::element_markdown (size = font$size, face = font$face, colour = font$color)
  }
  for(i in 1:length(list.plots)){
    p <- list.plots[[i]]
    if(is.ggplot(p)){
      if (!is.null(font.title)) p <- p + theme(plot.title = .set_font(font.title))
      if (!is.null(font.subtitle)) p <- p + theme(plot.subtitle = .set_font(font.subtitle))
      if (!is.null(font.caption)) p <- p + theme(plot.caption = .set_font(font.caption))
      if (!is.null(font.x)) p <- p + theme(axis.title.x = .set_font(font.x))
      if (!is.null(font.y)) p <- p + theme(axis.title.y = .set_font(font.y))
      if (!is.null(font.xtickslab)) p <- p + theme(axis.text.x = .set_font(font.xtickslab))
      if (!is.null(font.ytickslab)) p <- p + theme(axis.text.y = .set_font(font.ytickslab))
      list.plots[[i]] <- p
    }
  }
  if(is.ggplot(original.p)) list.plots[[1]]
  else list.plots
}

################### End of Helper Function #########################

#data: peri.final 
kaplan <- sh_cox
kaplan['m.time']<- kaplan$time/30

kaplan <- kaplan%>%
  relocate(m.time, .after=time)

# Model 1: NULL Model 
fit.kaplan <- survfit(Surv(m.time,event)~1, data = kaplan) 
#SDPID has a lot of missing data, SDPID.x/SDPID.y has no missing data

KM1<- ggsurvplot(fit.kaplan, 
           pval = TRUE, conf.int = TRUE, xlab = "Time (month)",
           risk.table = TRUE, #add risk table 
           break.time.by = 12, #break the time on risk table by 12 months
           risk.table.col = "strata", #change risk table colour by groups
           linetype = "strata", #change line type by groups
           surv.median.line = "hv", #specify median survival 
           ggtheme = theme_bw())

# Model 2: Sex
fit.kaplan.sex <- survfit(Surv(m.time,event)~ PtSex, data = kaplan) 

KM_sex <- ggsurvplot(fit.kaplan.sex, 
           size=1, #change line size
           legend.title = "Peritonitis-free interval by sex", legend.labs = c("Male","Female"), 
           pval = TRUE, conf.int = TRUE, xlab = "Time (month)",
           surv.plot.heigh =1,
           risk.table = TRUE, #add risk table 
           break.time.by = 12, #break the time on risk table by 12 months
           risk.table.col = "strata", #change risk table colour by groups
           risk.table.title="Number at risk of peritonitis", #change the risk table title 
           linetype = "strata", #change line type by groups
           surv.median.line = "hv", #specify median survival 
           ncensor.plot = TRUE, #show the number of censored patients 
           ggtheme = theme_bw())

#Customise the labels and font size 
KM_sex <- customize_labels(
  KM_sex,
  font.title    = c(8, "bold", "darkblue"),         
  font.subtitle = c(8, "bold.italic", "purple"), 
  font.caption  = c(8, "plain", "orange"),        
  font.x        = c(8, "bold.italic", "red"),          
  font.y        = c(8, "bold.italic", "darkred"),      
  font.xtickslab = c(8, "plain", "darkgreen")
)

# Model 3: Race 
fit.kaplan.race <- survfit(Surv(m.time,event)~ PtRace, data = kaplan) 

KM_race <- ggsurvplot(fit.kaplan.race, 
           size=1, #change line size
           legend.title = "Peritonitis-free interval by race", legend.labs = c("Malay","Chinese", "Indian", "Bumi.Sbh", "Bumi.Swk","Oth.Msian"), 
           pval = TRUE, conf.int = FALSE, xlab = "Time (month)",
           surv.plot.heigh =1,
           risk.table = TRUE, #add risk table 
           break.time.by = 12, #break the time on risk table by 12 months
           risk.table.col = "strata", #change risk table colour by groups
           risk.table.title="Number at risk of peritonitis", #change the risk table title 
           linetype = "strata", #change line type by groups
           surv.median.line = "hv", #specify median survival 
           ncensor.plot = FALSE, #show the number of censored patients 
           ggtheme = theme_bw())

# Model 4: Age group (KM only accepts categorical data) 
# two age groups was created: 55yo and above, 18-54 (why this cutoff?)
kaplan <- kaplan %>%
  mutate(Age2 = ifelse(Age >= 55, 1,0))

fit.kaplan.age <- survfit(Surv(m.time,event)~ Age2, data = kaplan) 

KM_age <- ggsurvplot(fit.kaplan.age, 
           size=1, #change line size
           legend.title = "Peritonitis-free interval by Age group", legend.labs = c("Below 55yo","Above 55yo"), 
           pval = TRUE, conf.int = FALSE, xlab = "Time (month)",
           surv.plot.heigh =1,
           risk.table = TRUE, #add risk table 
           break.time.by = 12, #break the time on risk table by 12 months
           risk.table.col = "strata", #change risk table colour by groups
           risk.table.title="Number at risk of peritonitis", #change the risk table title 
           linetype = "strata", #change line type by groups
           surv.median.line = "hv", #specify median survival 
           ncensor.plot = FALSE, #show the number of censored patients 
           ggtheme = theme_bw())

# Model 5: Diabetes as primary renal disease 
fit.kaplan.dm <- survfit(Surv(m.time,event)~ dm, data = kaplan) 

KM_dm <- ggsurvplot(fit.kaplan.dm, 
           size=1, #change line size
           legend.title = "Peritonitis-free interval by Primary Renal Disease (diabetes)", legend.labs = c("Not DM","Yes DM"), 
           pval = TRUE, conf.int = FALSE, xlab = "Time (month)",
           surv.plot.heigh =1,
           risk.table = TRUE, #add risk table 
           break.time.by = 12, #break the time on risk table by 12 months
           risk.table.col = "strata", #change risk table colour by groups
           risk.table.title="Number at risk of peritonitis", #change the risk table title 
           linetype = "strata", #change line type by groups
           surv.median.line = "hv", #specify median survival 
           ncensor.plot = FALSE, #show the number of censored patients 
           ggtheme = theme_bw())

# Model 6: Comorbidity of Cancer 
fit.kaplan.comob.cancer <- survfit(Surv(m.time,event)~ comob.cancer, data = kaplan) 

KM_comob.cancer <- ggsurvplot(fit.kaplan.comob.cancer, 
           size=1, #change line size
           legend.title = "Peritonitis-free interval by Cancer comorbidity", legend.labs = c("No cancer","Yes cancer"), 
           pval = TRUE, conf.int = FALSE, xlab = "Time (month)",
           surv.plot.heigh =1,
           risk.table = TRUE, #add risk table 
           break.time.by = 12, #break the time on risk table by 12 months
           risk.table.col = "strata", #change risk table colour by groups
           risk.table.title="Number at risk of peritonitis", #change the risk table title 
           linetype = "strata", #change line type by groups
           surv.median.line = "hv", #specify median survival 
           ncensor.plot = FALSE, #show the number of censored patients 
           ggtheme = theme_bw())

# Model 7: Comorbidity of CCF 
fit.kaplan.comob.ccf <- survfit(Surv(m.time,event)~ comob.ccf, data = kaplan) 

KM_comob.ccf <- ggsurvplot(fit.kaplan.comob.ccf, 
           size=1, #change line size
           legend.title = "Peritonitis-free interval by CCF comorbidity", legend.labs = c("No CCF","Yes CCF"), 
           pval = TRUE, conf.int = FALSE, xlab = "Time (month)",
           surv.plot.heigh =1,
           risk.table = TRUE, #add risk table 
           break.time.by = 12, #break the time on risk table by 12 months
           risk.table.col = "strata", #change risk table colour by groups
           risk.table.title="Number at risk of peritonitis", #change the risk table title 
           linetype = "strata", #change line type by groups
           surv.median.line = "hv", #specify median survival 
           ncensor.plot = FALSE, #show the number of censored patients 
           ggtheme = theme_bw())

# Model 8: Brand 
fit.kaplan.brand <- survfit(Surv(m.time,event)~ brand, data = kaplan) 

KM_brand <- ggsurvplot(fit.kaplan.brand, 
           size=1, #change line size
           legend.title = "Peritonitis-free interval by PD Brand", legend.labs = c("Baxter","Fresenius", "Lucenxia"), 
           pval = TRUE, conf.int = FALSE, xlab = "Time (month)",
           surv.plot.heigh =1,
           risk.table = TRUE, #add risk table 
           break.time.by = 12, #break the time on risk table by 12 months
           risk.table.col = "strata", #change risk table colour by groups
           risk.table.title="Number at risk of peritonitis", #change the risk table title 
           linetype = "strata", #change line type by groups
           surv.median.line = "hv", #specify median survival 
           ncensor.plot = FALSE, #show the number of censored patients 
           ggtheme = theme_bw())

summary(fit.kaplan)
summary(fit.kaplan)$table
summary(fit.kaplan, times=60)
summary(fit.kaplan, times=0:100)

``` 

```{r coxph, eval=T, echo=F, cache=T, results=FALSE, warning=FALSE, message=FALSE}

cox_sex<-coxph(Surv(m.time, event) ~ PtSex, data = kaplan)
summary(cox_sex)

#to conduct multiple UNIVARIATE cox regression with a helper function 
covariates<-c("Age","PtSex","toxic.nephropthy","glomerulonephritis", "obs.uropathy", "ADPKD", "day_diag", "day_pd", "hpt", "dm","comob.dm", "comob.hpt", "comob.peptic_ulcer", "comob.cerebrovas", "comob.chro_res", "comob.ihd", "comob.cancer", "comob.ccf", "comob.impair_vision", "comob.amputation", "comob.chron_liver","comob.peri_vas", "comob.psy", "comob.GI", "comob.cvd", "comob.tb","comob.ment_retard", "comob.dementia", "comob.sub_abuse")

# "PtRace"-> why this variable does not work? "PtAssist"

univ_formulas<- sapply(covariates, 
                       function(x) as.formula(paste('Surv(time, event)~', x)))

univ_models <- lapply(univ_formulas, function(x){coxph(x, data = kaplan)})
#Extract data
univ_results <- lapply(univ_models,
                       function(x){ 
                          x <- summary(x)
                          p.value<-signif(x$wald["pvalue"], digits=3)
                          wald.test<-signif(x$wald["test"], digits=3)
                          beta<-signif(x$coef[1], digits=3);#coeficient beta
                          HR <-signif(x$coef[2], digits=3);#exp(beta)
                          HR.confint.lower <- signif(x$conf.int[,"lower .95"], 2)
                          HR.confint.upper <- signif(x$conf.int[,"upper .95"],2)
                          HR <- paste0(HR, " (", 
                                       HR.confint.lower, "-", HR.confint.upper, ")")
                          res<-c(beta, HR, wald.test, p.value)
                          names(res)<-c("beta", "HR (95% CI for HR)", "wald.test", 
                                        "p.value")
                          return(res)
                          #return(exp(cbind(coef(x),confint(x))))
                         })
res <- t(as.data.frame(univ_results, check.names = FALSE))
as.data.frame(res)

# since PtRace does not work, we need to singularly test the variable
cox_race<-coxph(Surv(m.time, event) ~ PtRace, data = kaplan)
summary(cox_race)

#perhaps this is because some of the races were significant, while some were not significant -> this is why it causes many V1, V2, V3 to appear 

cox_assist<-coxph(Surv(m.time, event)~PtAssist, data=kaplan)
summary(cox_assist)

## Multivariate Cox regression analysis
multiv_cox <- coxph(Surv(m.time, event) ~ PtRace+Age+dm+comob.cancer+comob.ccf, data = kaplan)
summary(multiv_cox)

#Visualise the estimated distribution of survival times 
cox_plot <- ggsurvplot(survfit(multiv_cox), 
                       data = kaplan, 
                       palette= '#2E9FDF', 
                       break.time.by=12,
                       ggtheme = theme_minimal())

```

```{r cox_frailtypack, eval=F, echo=F, cache=T, results=FALSE, warning=FALSE, message=FALSE} 

cox.prop <- peri.final
cox.prop$time<- as.numeric(cox.prop$time)
cox.prop$day_diag<- as.numeric(cox.prop$day_diag)
cox.prop$day_pd<- as.numeric(cox.prop$day_pd)

class(cox.prop$time)

cox.prop<-rename(cox.prop, enum=Index)
cox.prop<-relocate(cox.prop, enum, .after=id)

library(frailtypack)
surv.cox<-with(cox.prop, Surv(time, event))

cox.model<- frailtyPenal(surv.cox~PtSex+PtRace+Age+comob.hpt+comob.dm+comob.peptic_ulcer+comob.other_cardiac+comob.cerebrovas+comob.chro_res+comob.ihd+comob.cancer+comob.ccf+comob.impair_vision+comob.chron_liver+comob.peri_vas+comob.GI+comob.cvd+comob.dementia+toxic.nephropthy+hpt+glomerulonephritis+dm+obs.uropathy+ADPKD, data=cox.prop, n.knots = 8, kappa = 1000)

summary(cox.model, level = 0.95)
plot(cox.model, type.plot = "Survival", level = 0.95, conf.bands = TRUE,
pos.legend = "topright", cex.legend = 0.7)


# for a shared model 
library(frailtypack)
surv.cox.shared<-with(cox.prop, Surv(time, event))

shared.model<- frailtyPenal(surv.cox.shared~cluster(PatientID)+PtSex+PtRace+Age+comob.hpt+comob.dm+comob.peptic_ulcer+comob.other_cardiac+comob.cerebrovas+comob.chro_res+comob.ihd+comob.cancer+comob.ccf+comob.impair_vision+comob.chron_liver+comob.peri_vas+comob.GI+comob.cvd+comob.dementia+toxic.nephropthy+hpt+glomerulonephritis+dm+obs.uropathy+ADPKD, data=cox.prop, n.knots = 8, kappa = 1000)

print(shared.model, digits = 4)
summary(shared.model, level = 0.95)
plot(shared.model, type.plot = "Survival", level = 0.95, conf.bands = TRUE,
pos.legend = "topright", cex.legend = 0.7)

```

##### End of Cox branch 


##### Fork to competing risk branch for peritonitis as outcome 
```{r wrangle_cmprsk_peri, eval=F, echo=F, cache=T, results=FALSE, warning=FALSE, message=FALSE}
# need to redo the steps in all_data EXCEPT dropping patients with PtOutcomeID==0
sh_final<-merge(sh7_8, sh_demo.fin, by="PatientID", all = TRUE)

sh_final<-arrange(sh_final, PatientID, PtDateOutcome)

## change column order 
##  | col_order = c("PatientID", "PtDateOutcome","PtOutcomeID", "event", "PtDateCommence", "PtDateCentre", "ESRFDateFirstRx", "day_diag", "day_pd") 

sh_final<-sh_final %>%
  relocate(PtDateCommence, .after=PatientID) %>%
  relocate(ESRFDateFirstRx, .after=PtDateCommence) %>%
  relocate(PtDateOutcome, .after = ESRFDateFirstRx) %>%
  relocate(PtOutcomeID, .after = PtDateOutcome) %>%
  relocate(day_diag, .after = PtOutcomeID) %>%
  relocate(day_pd, .after=day_diag) %>%
  relocate(event, .after=day_pd)

#If we don't drop PtOutcomeID==0, many rows will remain as NA

sh_final$PtDateOutcome[which(is.na(sh_final$PtDateOutcome))]<- as.Date("2018-12-31")
sh_final$event[which(is.na(sh_final$event))]<- 0 
sh_final$PtOutcomeID[which(is.na(sh_final$PtOutcomeID))]<- 0

# create t.start, tstop, time 
sh_final<-sh_final %>%
  mutate(t.start2=0) %>%
  relocate(t.start2, .after=event) %>%
  mutate(t.stop=PtDateOutcome - PtDateCommence) %>%
  relocate(t.stop, .after=t.start2)

sh_final<-sh_final %>%
  group_by(PatientID) %>%
  mutate(Index=row_number()) %>%
  relocate(Index, .after=t.stop) 

sh_final<-sh_final %>%
  group_by(PatientID) %>%
  mutate(t.start = ifelse(Index>1, lag(t.stop), 0)) %>%
  relocate(t.start, .before=t.stop)
  
 ## sh_final$t.start[sh_final$Index>1]<-lead(sh_final$t.stop) #this doesn't work 

## create a third duration variable 'time' 
sh_final['time']<-sh_final$t.stop - sh_final$t.start
sh_final<-relocate(sh_final, time, .after = t.stop)



sh_final <- sh_final %>%
  relocate(PtSex, .after=Index) %>%
  relocate(PtRace, .after=PtSex) %>%
  relocate(Age, .after=PtRace)

sh_final <- sh_final %>%
  mutate(death=ifelse(PtOutcomeID==1, 1, 0)) %>%
  relocate(death, .after=Age)

# check the category of Patient Outcome 
plot_bar(sh_final$PtOutcomeID)  #shows 7 categories
                                #12: peritonitis, 1: death, 13: ESI, 
                                #2: HD, 8: recover kidney function,                                 
                                #6: lost to f/up, 3: transplanted 

comp <- sh_final #prepare a new datasets for wrangling for cox

#create a new variable "status" 
comp['status']<- NA
comp <- comp %>% 
  relocate(status, .after= PtOutcomeID)

comp <- comp %>% 
  arrange(PatientID, PtDateOutcome) %>%
  mutate(Index=row_number()) %>%
  relocate(Index, .after=status) 

#find out which subject has index==1 and outcome==0 -> need to remove specifically these rows
comp2<-comp %>% 
  filter(n()!=1)

which(comp2$Index==1 & comp2$PtOutcomeID ==0)   #no subject had outcome==0 as first outcome event 
                                                ## comp2 has no more value 

comp <- comp %>% 
  arrange(PatientID, PtDateOutcome)  %>%
  filter(Index==1)

## | use the newer pivot_wider() from tidyverse/tidyr to make the conversion 
## | add a new column to indicate primary renal disease is "present" 
#comp['prim.dis']<-as.integer(1)
#class(comp$prim.dis)

#comp <- comp %>%
#  pivot_wider(
#    names_from = PtRenal1, 
#    values_from = prim.dis
#  )

##  | rename the new columns 
#names(comp)[64:74] <- c("toxic.nephropthy", "hpt","glomerulonephritis", "dm", "unknown", "obs.uropathy","nil", "miss","others", "hereditary.nephritis", "ADPKD")

##  | drop column nil and miss 
#comp<- comp[,-(70:71)]

##  | drop PtHeight and PtWeight (for now since too many missing data)
comp <- subset(comp, select = -c(t.start2, NotifID, PtRRTID, SDPID, SDPID.x, PtARID, NotifID.x, PtRRTID.x, givendate.x,NotifID.y, PtRRTID.y, PtRenalSpecify, PtRenal2, PtRenalSpecify2, givendate.y, PtHeight, PtWeight, Modality))

#fill in "status": 0=censor, 1=peritonitis, 2=competing risk events 
comp$status[which(comp$PtOutcomeID==0)]<-0
comp$status[which(comp$PtOutcomeID==12)]<-1
comp$status[which(is.na(comp$status))]<-2


#expand the variable "PtSex" and "PtRace" to dummy tables 
#comp['female'] <- NA
#comp$female[which(comp$PtSex==2)]<-1
#comp$female[which(is.na(comp$female))]<-0 

#comp['race_ind']<-as.integer(1)

#comp <- comp %>%
#  pivot_wider(
#    names_from = PtRace, 
#    values_from = race_ind
#  )

#names(comp)[55:60] <- c("malay", "chinese","indian", "bumi.sarawak", "others", "bumi.sabah")

comp[is.na(comp)]<-0

#convert day_diag and day_pd to months 
comp['month_diag']<-comp$day_diag/30
comp['month_pd']<- comp$day_pd/30

######## Convert SDPID.y to Site (arranged by State) ######## 
comp['site']<- NA
comp$site[which(comp$SDPID.y==330)]<-2
comp$site[which(comp$SDPID.y==430)]<-3
comp$site[which(comp$SDPID.y==530 | comp$SDPID.y==630)]<-4
comp$site[which(comp$SDPID.y==830 | comp$SDPID.y==831)]<-5
comp$site[which(comp$SDPID.y==730 | comp$SDPID.y==4330 | comp$SDPID.y==108130)]<-6
comp$site[which(comp$SDPID.y==930)]<-7
comp$site[which(comp$SDPID.y==1030 | comp$SDPID.y==2330 | comp$SDPID.y==3030 | comp$SDPID.y==7731)]<-8
comp$site[which(comp$SDPID.y==1330)]<-9
comp$site[which(comp$SDPID.y==1230 | comp$SDPID.y==3130 | comp$SDPID.y==3730 | comp$SDPID.y==5030 | comp$SDPID.y==7130 | comp$SDPID.y==143730)]<-10
comp$site[which(comp$SDPID.y==1130 | comp$SDPID.y==8530 | comp$SDPID.y==106530)]<-11
comp$site[which(comp$SDPID.y==1430 | comp$SDPID.y==1530 | comp$SDPID.y==1630)]<-12
comp$site[which(comp$SDPID.y==1730 | comp$SDPID.y==1930)]<-13
comp$site[which(comp$SDPID.y==130)]<-14

comp$site <- as.factor(comp$site)

## combine the dataset with sh11_pdsys 
comp <- merge(comp, sh11_pdsys, by="PatientID", all=TRUE)

```

```{r factor2ind, eval=F, echo=F, cache=T, results=FALSE}
library(cmprsk)

##############################################################################
# Adds-on functions to crr() function in 'cmprsk' package by Gray, RJ. 
# Written by Luca Scrucca
#
# Reference: 
# Scrucca L, Santucci A, Aversa F (2009) Regression Modeling of Competing 
#   Risk Using R: An In Depth Guide for Clinicians. Submitted to Bone Marrow
#   Transplantation
##############################################################################

# This ensure that the package is loaded
if(!require(cmprsk))
  { stop("the package 'cmprsk' is required, please install it. \nSee help(install.packages).") }
  
factor2ind <- function(x, baseline)
{
# Given a factor variable x, create an indicator matrix of dimension 
# length(x) x (nlevels(x)-1) dropping the column corresponding to the 
# baseline level (by default the first level is used as baseline).
# Example:
# > x = gl(4, 2, labels = LETTERS[1:4])
# > factor2ind(x)
# > factor2ind(x, "C")
  xname <- deparse(substitute(x))
  n <- length(x)
  x <- as.factor(x)
  if(!missing(baseline)) x <- relevel(x, baseline)
  X <- matrix(0, n, length(levels(x)))
  X[(1:n) + n*(unclass(x)-1)] <- 1
  X[is.na(x),] <- NA
  dimnames(X) <- list(names(x), paste(xname, levels(x), sep = ":"))
  return(X[,-1,drop=FALSE])
}

modsel.crr <- function (object, ..., d = log(object$n)) 
{
  if(class(object) != "crr") 
    stop("object is not of class 'crr'")
  objects <- list(object, ...)
  nmodels <- length(objects)
  modnames <- paste("Model ", format(1:nmodels), ": ", 
                      lapply(objects, function(x) x$call), 
                      sep = "", collapse = "\n")
  # add null model
  mod0 <- object
  mod0$loglik <- mod0$loglik.null
  mod0$coef <- mod0$call$cov1 <- mod0$call$cov2 <- NULL
  objects <- c(list(mod0), objects)
  nmodels <- nmodels + 1
  #
  modnames <- c("Model 0: Null model", modnames)
  ns <- sapply(objects, function(x) x$n) 
  dfs <- sapply(objects, function(x) length(x$coef)) 
  if(any(ns != ns[1]))
    stop("models were not all fitted to the same dataset")
  out <- matrix(rep(NA, 5 * nmodels), ncol = 5)
  loglik <- sapply(objects, function(x) x$loglik)
  crit <- sapply(objects, function(x) -2*x$loglik + d*length(x$coef))
  out[,1] <- ns
  out[,2] <- loglik
  out[,3] <- dfs
  out[,4] <- crit
  out[,5] <- crit - min(crit)
  if(d==log(object$n)) critname <- "BIC"
  else if(d == 2) critname <- "AIC"
  else critname <- "Criterion"
  colnames(out) <- c("Num.obs", "logLik", "Df.fit", critname, paste(critname, "diff"))
  rownames(out) <- 0:(nmodels-1)
  title <- "Model selection table\n"
  topnote <- modnames
  structure(as.data.frame(out), heading = c(title, topnote), 
            class = c("anova", "data.frame"))
}



factor2ind(comp$PtSex, 2) #female = 2
factor2ind(comp$PtRace, 1)
factor2ind(comp$SDPID.y, "130") #Unknokwn site - find out; this is "factor" class but why "130" as string? 
factor2ind(comp$PtRenal1, 2) #Diabetes =2 
factor2ind(comp$PtAssist, 1) #complete self-care=1
factor2ind(comp$site, 2) #Kedah = 2
factor2ind(comp$PDRegimen, "1") #CAPD =1 
factor2ind(comp$PDSystem, "101") #UltraBag = 101
factor2ind(comp$brand, "1")#Baxter=1


```

```{r cmprsk_peri, eval=F, echo=F, cache=T, results=FALSE, warning=FALSE, message=FALSE}

#x <- subset(comp, select = c(Age, month_diag, month_pd, (22:42), (45:60) ))

x <- cbind(comp$Age, factor2ind(comp$PtSex, 2),factor2ind(comp$PtRace, 1),factor2ind(comp$PtRenal1, 1),factor2ind(comp$PtAssist, 1),factor2ind(comp$site, 2), comp$month_pd, comp$month_diag, factor2ind(comp$brand, 1))
head(x)

#x <- cbind(factor2ind(comp$PtRace, 1),factor2ind(comp$PtRenal1, 1),factor2ind(comp$PtAssist, 1),factor2ind(comp$SDPID.y, "130"))

str(comp)
comp$time <- as.numeric(comp$time)
comp$month_diag<-as.numeric(comp$month_diag)
comp$month_pd<-as.numeric(comp$month_pd)

library(cmprsk)
mod1<- crr(ftime = comp$time, fstatus = comp$status, x, failcode = 1, cencode = 0) #status=1, peritonitis 

#mod1<- crr(ftime = comp$time, fstatus = comp$status, x, failcode = 2, cencode = 0)

#mod1<- crr(ftime = comp$time, fstatus = comp$status, x)

summary(mod1, digits = 4)

#The overall p-value is necessary when modeling a factor with more than two levels; Wald test can be done to obtained this p-value 

library(aod)
wald.test(mod1$var, mod1$coef, Terms = 3:7) #p-value for race 
wald.test(mod1$var, mod1$coef, Terms = 8:15) #primary renal disease, but i ignore last two levels 
wald.test(mod1$var, mod1$coef, Terms = 18:20)

#model selection 
mod2<- crr(ftime = comp$time, fstatus = comp$status, x[,3:7]) #race
mod3<- crr(ftime = comp$time, fstatus = comp$status, x[,8:15]) #Prim.Renal Disease
mod4<- crr(ftime = comp$time, fstatus = comp$status, x[,18:20])#Assistance 
mod5<- crr(ftime = comp$time, fstatus = comp$status, x[,22:33]) #site
mod6<- crr(ftime = comp$time, fstatus = comp$status, x[,36:37]) #brand

mod7<- crr(ftime = comp$time, fstatus = comp$status, x[,c(3:7,8:15)]) #race, Renal1
mod8<- crr(ftime = comp$time, fstatus = comp$status, x[,c(3:7,18:20)]) #race, Assistance
mod9<- crr(ftime = comp$time, fstatus = comp$status, x[,c(3:7,22:33)]) #race, Site
mod10<- crr(ftime = comp$time, fstatus = comp$status, x[,c(3:7,36:37)]) #race, brand
mod11<- crr(ftime = comp$time, fstatus = comp$status, x[,c(8:15, 18:20)]) #Renal1, Assist
mod12<- crr(ftime = comp$time, fstatus = comp$status, x[,c(8:15, 22:33)]) #rRenal1, site
mod13<- crr(ftime = comp$time, fstatus = comp$status, x[,c(8:15, 36:37)]) #Renal1, brand
mod14<- crr(ftime = comp$time, fstatus = comp$status, x[,c(22:33, 36:37)]) #Site, Brand 

mod15<- crr(ftime = comp$time, fstatus = comp$status, x[,c(3:7,8:15,18:20)])#race, Renal1, assist
mod16<- crr(ftime = comp$time, fstatus = comp$status, x[,c(3:7,8:15,22:33)])#race, renal1, site 
mod17<- crr(ftime = comp$time, fstatus = comp$status, x[,c(3:7,8:15,36:37)])#race, renal1, brand 
mod18<- crr(ftime = comp$time, fstatus = comp$status, x[,c(8:15,18:20,22:33)])#renal1, assist, site 
mod19<- crr(ftime = comp$time, fstatus = comp$status, x[,c(8:15,18:20, 36:37)])#renal1, assist, brand 
mod20<- crr(ftime = comp$time, fstatus = comp$status, x[,c(18:20, 22:33, 36:37)])#assist, site, brand 
mod21<- crr(ftime = comp$time, fstatus = comp$status, x[,c(3:7, 22:33, 36:37)])#Race, site, brand 

mod22<- crr(ftime = comp$time, fstatus = comp$status, x[,c(3:7,8:15,18:20, 22:33)]) #race, renal1, assist, site 
mod23<- crr(ftime = comp$time, fstatus = comp$status, x[,c(3:7,8:15,18:20, 36:37)]) #race, renal1, assist, brand 

mod24<- crr(ftime = comp$time, fstatus = comp$status, x[,c(3:7,8:15,18:20, 22:33, 36:37)]) #all 


modsel.crr(mod1, mod2, mod3, mod4, mod5, mod6, mod7, mod8, mod9, mod10, mod11, mod12, mod13, mod14, mod15, mod16, mod17, mod18, mod19, mod20, mod21, mod22, mod23, mod24)

summary(mod6)

#Model diagnostic 
mod2$res
par(mfrow=c(1,3), mar=c(4.5,4,2,1))
for(j in 1:ncol(mod2$res))
  scatter.smooth(mod2$uft, mod2$res[,j],
                 main = names(mod2$coef)[j],
                 xlab = "Failure time",
                 ylab = "Schoenfeld residuals")

# need to read more about Schoenfield? plot 

# Model-based estimation of the CIF 
x0 = cbind(comp$PtRace, factor2ind(levels(comp$PtRace), "1"))
x0  

pred<-predict(mod2, x0)
pred

plot(pred, lty = 1:4, xlab="failure time", ylab="CIF")
legend("topleft", legend = levels(comp$PtRace), lty = 1:6, title = "Race")


cuminc(ftime = comp$time, fstatus = comp$status, cencode = 0, na.action = na.omit )

```

##### End of the Fork to competing risk branch for peritonitis 



##### Fork to competing risk branch for "transfer to HD" as outcome 
```{r wrangle_cmprsk_peri, eval=F, echo=F, cache=T, results=FALSE, warning=FALSE, message=FALSE}

sh_final.hd<-merge(sh7_outcome, sh_demo.fin, by="PatientID", all = TRUE)

sh_final.hd<-arrange(sh_final.hd, PatientID, PtDateOutcome)

# create a new column for "event" 
sh_final.hd<-sh_final.hd %>%
  mutate(event=ifelse(PtOutcomeID==2, 1,0)) %>%
  relocate(event, .after=PtOutcomeID)

## change column order 
##  | col_order = c("PatientID", "PtDateOutcome","PtOutcomeID", "event", "PtDateCommence", "PtDateCentre", "ESRFDateFirstRx", "day_diag", "day_pd") 

sh_final.hd<-sh_final.hd %>%
  relocate(PtDateCommence, .after=PatientID) %>%
  relocate(ESRFDateFirstRx, .after=PtDateCommence) %>%
  relocate(PtDateOutcome, .after = ESRFDateFirstRx) %>%
  relocate(PtOutcomeID, .after = PtDateOutcome) %>%
  relocate(day_diag, .after = PtOutcomeID) %>%
  relocate(day_pd, .after=day_diag) %>%
  relocate(event, .after=day_pd)

#If we don't drop PtOutcomeID==0, many rows will remain as NA

sh_final.hd$PtDateOutcome[which(is.na(sh_final.hd$PtDateOutcome))]<- as.Date("2018-12-31")
sh_final.hd$event[which(is.na(sh_final.hd$event))]<- 0 
sh_final.hd$PtOutcomeID[which(is.na(sh_final.hd$PtOutcomeID))]<- 0

# create t.start, tstop, time 
sh_final.hd<-sh_final.hd %>%
  mutate(t.start2=0) %>%
  relocate(t.start2, .after=event) %>%
  mutate(t.stop=PtDateOutcome - PtDateCommence) %>%
  relocate(t.stop, .after=t.start2)

sh_final.hd<-sh_final.hd %>%
  group_by(PatientID) %>%
  mutate(Index=row_number()) %>%
  relocate(Index, .after=t.stop) 

sh_final.hd<-sh_final.hd %>%
  group_by(PatientID) %>%
  mutate(t.start = ifelse(Index>1, lag(t.stop), 0)) %>%
  relocate(t.start, .before=t.stop)
  
 ## sh_final$t.start[sh_final$Index>1]<-lead(sh_final$t.stop) #this doesn't work 

## create a third duration variable 'time' 
sh_final.hd['time']<-sh_final.hd$t.stop - sh_final.hd$t.start
sh_final.hd<-relocate(sh_final.hd, time, .after = t.stop)



sh_final.hd <- sh_final.hd %>%
  relocate(PtSex, .after=Index) %>%
  relocate(PtRace, .after=PtSex) %>%
  relocate(Age, .after=PtRace)

sh_final.hd <- sh_final.hd %>%
  mutate(death=ifelse(PtOutcomeID==1, 1, 0)) %>%
  relocate(death, .after=Age)

# check the category of Patient Outcome 
plot_bar(sh_final.hd$PtOutcomeID)  #shows 6 categories
                                #1: death, 0: censor
                                #2: HD, 8: recover kidney function,                                 
                                #6: lost to f/up, 3: transplanted 

comp_hd <- sh_final.hd #prepare a new datasets for wrangling for cox

#create a new variable "status" 
comp_hd['status']<- NA
comp_hd <- comp_hd %>% 
  relocate(status, .after= PtOutcomeID)

comp_hd <- comp_hd %>% 
  arrange(PatientID, PtDateOutcome) %>%
  mutate(Index=row_number()) %>%
  relocate(Index, .after=status) 

#find out which subject has index==1 and outcome==0 -> need to remove specifically these rows
comp2_hd<-comp_hd %>% 
  filter(n()!=1)

which(comp2_hd$Index==1 & comp2_hd$PtOutcomeID ==0) 

#set a new dataset for HD as outcome (technique failure)
view(comp_hd)

##  | drop PtHeight and PtWeight (for now since too many missing data)
comp_hd <- subset(comp_hd, select = -c(t.start2, NotifID, PtRRTID, SDPID, SDPID.x, PtARID, NotifID.x, PtRRTID.x, givendate.x,NotifID.y, PtRRTID.y, PtRenalSpecify, PtRenal2, PtRenalSpecify2, givendate.y, PtHeight, PtWeight, Modality, PtNotifOutcomeID, PtDateOutcomeNotified))


comp_hd <- arrange(comp_hd, PatientID, PtDateOutcome)

#create a new column named "unique" to identify unique patients 
comp_hd['uniq']<-NA

comp_hd<- comp_hd %>%
  group_by(PatientID) %>%
  mutate(uniq = ifelse(n()==1,1, 0)) %>%
  relocate(uniq, .after= "Index")

comp_hd<- comp_hd %>% 
  group_by(PatientID) %>%
  mutate(uniq2 = ifelse(duplicated(PtOutcomeID) & uniq==0, 1, 0)) %>%   #identify duplicated outcome 
  relocate(uniq2, .after="uniq")                                        #without deleting the ONLY entry as 
                                                                        #"duplicated" will label the first entry as "0"

which(comp_hd$uniq2==1) #index 401 (ID 70679), 533 (ID 71571)
#ID 70679 PtOutcomeID has two "transfer to HD" at different dates - Sol: remove the second HD episode (probably it signified a transfer back to PD)
#ID 71571 Patient has two "transfer to HD" records on 2016-08-08 and 2017-02-01 - Sol: remove the second "transfer to HD" entry

comp_hd<-comp_hd[!(comp_hd$PatientID==70679 & comp_hd$PtDateOutcome == "2017-05-01"), ]
comp_hd<-comp_hd[!(comp_hd$PatientID==71571 & comp_hd$PtDateOutcome == "2017-02-01"), ]

comp_hd<- comp_hd %>%
  group_by(PatientID) %>%
  mutate(uniq3 = ifelse(n()==1,1, 0)) %>%
  relocate(uniq3, .after= "uniq2")  

#ID 70788 Patient died first (2017-05-14) before being "lost to follow up" on 2017-09-26 - Sol: remove the "lost to follow-up" entry 
#ID 70787 Patient "lost to follow up" on 2017-05-30 then "died" on 2017-9-26 - Sol: remove the "lost to follow-up entry"
#ID 71495 Patient "lost to follow up" on 2015-01-01 then "died" on 2015-02-23 - Sol: remove the "lost to follow-up entry"
#ID 75599 Patient "lost to follow up" on 2017-04-02 then "died" on 2017-04-14 - Sol: remove the "lost to follow-up entry"
#ID 76102 Patient "lost to follow up" on 2016-01-09 then "died" on 2016-01-10 - Sol: remove the "lost to follow-up entry" 

comp_hd<-comp_hd[!(comp_hd$PatientID==70788 & comp_hd$PtDateOutcome == "2017-12-01"), ]
comp_hd<-comp_hd[!(comp_hd$PatientID==70787 & comp_hd$PtDateOutcome == "2017-05-30"), ]
comp_hd<-comp_hd[!(comp_hd$PatientID==71495 & comp_hd$PtDateOutcome == "2015-01-01"), ]
comp_hd<-comp_hd[!(comp_hd$PatientID==75599 & comp_hd$PtDateOutcome == "2017-04-02"), ]
comp_hd<-comp_hd[!(comp_hd$PatientID==76102 & comp_hd$PtDateOutcome == "2016-01-09"), ]

comp_hd<- comp_hd %>%
  group_by(PatientID) %>%
  mutate(uniq4 = ifelse(n()==1,1, 0)) %>%
  relocate(uniq4, .after= "uniq3")

comp_hd<- comp_hd[!(comp_hd$PtOutcomeID==1 & comp_hd$uniq4==0) ,]

#fill in "status": 0=censor, 1=technique failure (transfer to HD), 2=competing risk events (death)
comp_hd$status[which(comp_hd$PtOutcomeID==0)]<-0
comp_hd$status[which(comp_hd$PtOutcomeID==2)]<-1 
comp_hd$status[which(is.na(comp_hd$status))]<-2

comp_hd[is.na(comp_hd)]<-0







#convert day_diag and day_pd to months 
comp_hd['month_diag']<-comp_hd$day_diag/30
comp_hd['month_pd']<- comp_hd$day_pd/30

######## Convert SDPID.y to Site (arranged by State) ######## 
comp_hd['site']<- NA
comp_hd$site[which(comp_hd$SDPID.y==330)]<-2
comp_hd$site[which(comp_hd$SDPID.y==430)]<-3
comp_hd$site[which(comp_hd$SDPID.y==530 | comp_hd$SDPID.y==630)]<-4
comp_hd$site[which(comp_hd$SDPID.y==830 | comp_hd$SDPID.y==831)]<-5
comp_hd$site[which(comp_hd$SDPID.y==730 | comp_hd$SDPID.y==4330 | comp_hd$SDPID.y==108130)]<-6
comp_hd$site[which(comp_hd$SDPID.y==930)]<-7
comp_hd$site[which(comp_hd$SDPID.y==1030 | comp_hd$SDPID.y==2330 | comp_hd$SDPID.y==3030 | comp_hd$SDPID.y==7731)]<-8
comp_hd$site[which(comp_hd$SDPID.y==1330)]<-9
comp_hd$site[which(comp_hd$SDPID.y==1230 | comp_hd$SDPID.y==3130 | comp_hd$SDPID.y==3730 | comp_hd$SDPID.y==5030 | comp_hd$SDPID.y==7130 | comp_hd$SDPID.y==143730)]<-10
comp_hd$site[which(comp_hd$SDPID.y==1130 | comp_hd$SDPID.y==8530 | comp_hd$SDPID.y==106530)]<-11
comp_hd$site[which(comp_hd$SDPID.y==1430 | comp_hd$SDPID.y==1530 | comp_hd$SDPID.y==1630)]<-12
comp_hd$site[which(comp_hd$SDPID.y==1730 | comp_hd$SDPID.y==1930)]<-13
comp_hd$site[which(comp_hd$SDPID.y==130)]<-14

comp_hd$site <- as.factor(comp_hd$site)

## combine the dataset with sh11_pdsys 
comp_hd <- merge(comp_hd, sh11_pdsys, by="PatientID", all=TRUE)

##################
### factor2ind ###
##################

factor2ind(comp_hd$PtSex, 2) #female = 2
factor2ind(comp_hd$PtRace, 1)
factor2ind(comp_hd$SDPID.y, "130") #Unknokwn site - find out; this is "factor" class but why "130" as string? 
factor2ind(comp_hd$PtRenal1, 2) #Diabetes =2 
factor2ind(comp_hd$PtAssist, 1) #complete self-care=1

factor2ind(comp_hd$site, 2) #Kedah = 2
factor2ind(comp_hd$PDRegimen, "1") #CAPD =1 
factor2ind(comp_hd$PDSystem, "101") #UltraBag = 101
factor2ind(comp_hd$brand, "1")#Baxter=1

```

```{r cmprsk_hd, eval=F, echo=F, cache=T, results=FALSE, warning=FALSE, message=FALSE}

x <- cbind(comp_hd$Age, factor2ind(comp_hd$PtSex, 2),factor2ind(comp_hd$PtRace, 1),factor2ind(comp_hd$PtRenal1, 1),factor2ind(comp_hd$PtAssist, 1),factor2ind(comp_hd$site, 2), comp_hd$month_pd, comp_hd$month_diag, factor2ind(comp_hd$brand, 1))
head(x)


str(comp_hd)
comp_hd$time <- as.numeric(comp_hd$time)
comp_hd$month_diag<-as.numeric(comp_hd$month_diag)
comp_hd$month_pd<-as.numeric(comp_hd$month_pd)

comp_hd['m.time']<- as.numeric(comp_hd$time/30)

library(cmprsk)
mod1<- crr(ftime = comp_hd$m.time, fstatus = comp_hd$status, x, failcode = 1, cencode = 0) #status=1, HD/technique failure 

#mod1<- crr(ftime = comp_hd$time, fstatus = comp_hd$status, x, failcode = 2, cencode = 0)

#mod1<- crr(ftime = comp_hd$time, fstatus = comp_hd$status, x)

summary(mod1, digits = 4)

#The overall p-value is necessary when modeling a factor with more than two levels; Wald test can be done to obtained this p-value 

#Variables that are significant (to be included in subsequent model selection), with p<0.1
#Age, Renal1 (8:15), Assist (18:20), site (22:33)

library(aod)
wald.test(mod1$var, mod1$coef, Terms = 1) #p-value for age
wald.test(mod1$var, mod1$coef, Terms = 8:15) #primary renal disease, but i ignore last two levels 
wald.test(mod1$var, mod1$coef, Terms = 18:20) #assist 
wald.test(mod1$var, mod1$coef, Terms = 22:33) #site


#model selection 
mod2<- crr(ftime = comp_hd$m.time, fstatus = comp_hd$status, x[,1]) #age
mod3<- crr(ftime = comp_hd$m.time, fstatus = comp_hd$status, x[,8:15]) #Prim.Renal Disease
mod4<- crr(ftime = comp_hd$m.time, fstatus = comp_hd$status, x[,18:20])#Assistance 
mod5<- crr(ftime = comp_hd$m.time, fstatus = comp_hd$status, x[,22:33]) #site

mod6<- crr(ftime = comp_hd$m.time, fstatus = comp_hd$status, x[,c(1,8:15)]) #age, Renal1
mod7<- crr(ftime = comp_hd$m.time, fstatus = comp_hd$status, x[,c(1,18:20)]) #age, Assistance
mod8<- crr(ftime = comp_hd$m.time, fstatus = comp_hd$status, x[,c(1,22:33)]) #age, Site

mod9<- crr(ftime = comp_hd$m.time, fstatus = comp_hd$status, x[,c(8:15, 18:20)]) #Renal1, Assist
mod10<- crr(ftime = comp_hd$m.time, fstatus = comp_hd$status, x[,c(8:15, 22:33)]) #rRenal1, site
mod11<- crr(ftime = comp_hd$m.time, fstatus = comp_hd$status, x[,c(18:20, 22:33)]) #assist, site

mod12<- crr(ftime = comp_hd$m.time, fstatus = comp_hd$status, x[,c(1,8:15,18:20)])#age, Renal1, assist
mod13<- crr(ftime = comp_hd$m.time, fstatus = comp_hd$status, x[,c(1,8:15,22:33)])#age, renal1, site 
mod14<- crr(ftime = comp_hd$m.time, fstatus = comp_hd$status, x[,c(8:15,18:20,22:33)])#renal1, assist, site 
mod15<- crr(ftime = comp_hd$m.time, fstatus = comp_hd$status, x[,c(1, 18:20, 22:33)])#age, assist, site 

mod16<- crr(ftime = comp_hd$m.time, fstatus = comp_hd$status, x[,c(1, 8:15, 18:20, 22:33)])#age, renal1, assist, site 

modsel.crr(mod1, mod2, mod3, mod4, mod5, mod6, mod7, mod8, mod9, mod10, mod11, mod12, mod13, mod14, mod15, mod16)



summary(mod7)

#Model diagnostic 
mod7$res
par(mfrow=c(1,3), mar=c(4.5,4,2,1))
for(j in 1:ncol(mod7$res))
  scatter.smooth(mod7$uft, mod7$res[,j],
                 main = names(mod7$coef)[j],
                 xlab = "Failure time",
                 ylab = "Schoenfeld residuals")

# need to read more about Schoenfield? plot 

# Model-based estimation of the CIF 
x0 = cbind(comp_hd$PtAssist, factor2ind(levels(comp_hd$PtAssist), 1))
x0  

pred<-predict(mod7, x0)
pred

plot(pred, lty = 1:4, xlab="failure time", ylab="CIF")
legend("topleft", legend = levels(comp_hd$PtAssist), lty = 1:4, title = "Assistance for CAPD")


cuminc(ftime = comp_hd$m.time, fstatus = comp_hd$status, cencode = 0, na.action = na.omit )


###############################
### use the crrstep package ### <- not working 
###############################
library(crrstep)

comp_hd['ftime']<-comp_hd$m.time
comp_hd$ftime<-as.numeric(comp_hd$ftime)

#x <- cbind(comp_hd$Age, factor2ind(comp_hd$PtSex, 2),factor2ind(comp_hd$PtRace, 1),factor2ind(comp_hd$PtRenal1, 1),factor2ind(comp_hd$PtAssist, 1),factor2ind(comp_hd$site, 2), comp_hd$month_pd, comp_hd$month_diag, factor2ind(comp_hd$brand, 1))

formula1<- comp_hd$ftime~1+comp_hd$Age+as.factor(comp_hd$PtSex)+as.factor(comp_hd$PtRace)+as.factor(comp_hd$PtRenal1)

crrstep(formula1, scope.min = ~1,etype = 1, fstatus=comp_hd$status, data=comp_hd, direction="forward",failcode=1,cencode=0, criterion="BIC")


###########################
### fastcmprisk package ### <-not working yet 
###########################

library(crrp)
lam.path <- 10^seq(log10(0.1), log10(0.001), length = 25)

library(cmprsk)
mod1<- crr(ftime = comp_hd$m.time, fstatus = comp_hd$status, x, failcode = 1, cencode = 0) #status=1, HD/technique failure 

mod1.fast <- fastcmprsk::fastCrr(comp_hd$m.time, comp_hd$status, x, variance=FALSE, returnDataFrame=TRUE)

library(fastcmprsk)

 set.seed(2019)
nobs <- 500
beta1 <- c(0.40, -0.40, 0, -0.50, 0, 0.60, 0.75, 0, 0, -0.80)
beta2 <- -beta1
Z <- matrix(rnorm(nobs * length(beta1)), nrow = nobs)
dat <- simulateTwoCauseFineGrayModel(nobs, beta1, beta2, Z, u.min = 0, u.max = 1, p = 0.5)

```
##### End of competing risk branch HD





```{r wrangling, eval=F, echo=F, cache=T, results=FALSE, warning=FALSE, message=FALSE} 

# | Types of analysis we can do with this database 

# Cox proportional hazard model 
## Time to first peritonitis 

# Shared frailty model 
## Obs clustered into groups - one layer of clusters 
##  | recurrent event, hospitals or cities, PD products? 

# Nested frailty model 
## Hierarchicl clustering - includes two nested random effects that act multiplicatively on hazard function 
##  | hospitals nested within state, different PD products nested within company? 

# Joint frailty model 
## Recurrent events that are terminated by loss to f/up, end-of-study, or major failure (death)
##  | recurrent peritonitis terminated by HD, or death 

# Compulsory variables: PatientID, t.start, t.stop, time, event

# sh1: PatientID, PtDateBirth, PtSex, PtCitizenship, PtRace, PtABO, Age, [givendate]
# sh3: [PtARID], PatientID, [NotifID],  [PtRRTID], SDPID, DataYear, PtAssist, Asst.chang, deteriorate 
# sh4: PatientID, [NotifID], [PtRRTID], SDPID, [Modality], PtDateCommenc, PtDateCentre, ESRFDateFirstRx, PtHeight, PtWeight, PtRenal1, [PtRenalSpecify],[PtRenal2], [PtRenalSpecify2], day_diag, day_pd, [givendate]
# sh5: comorbidity - include all columns 
# sh6: lab results - ignore all for now 


# create a shorter data.frame sh_final2 
sh_final <- sh_final %>%
  relocate(PtSex, .after=Index) %>%
  relocate(PtRace, .after=PtSex) %>%
  relocate(Age, .after=PtRace)

sh_final <- sh_final %>%
  mutate(death=ifelse(PtOutcomeID==1, 1, 0)) %>%
  relocate(death, .after=Age)

## From here, need to customise the dataset according to the types of analysis 
  sh_final2<- sh_final[, c(1:17)]
  sh_final2<-subset(sh_final2, select = -c(t.start2, PtDateOutcome, PtDateCommence, ESRFDateFirstRx))

sh_final2<- sh_final2 %>%
  relocate(t.start, .after=PatientID) %>%
  relocate(t.stop, .after=t.start) %>%
  relocate(time, .after=t.stop) %>%
  relocate(event, .after=time) %>%
  rename(id=PatientID)

sh_final2$id<-as.integer(sh_final2$id) 
sh_final2$t.start<-as.integer(sh_final2$t.start)
sh_final2$t.stop<-as.integer(sh_final2$t.stop)
sh_final2$time<-as.integer(sh_final2$time)
sh_final2$event<-as.integer(sh_final2$event)
sh_final2$PtSex<-as.factor(sh_final2$PtSex)
sh_final2$PtRace<-as.factor(sh_final2$PtRace)
sh_final2$Age<-as.numeric(sh_final2$Age)

sh_final2<-rename(sh_final2, enum=Index)
sh_final2<-relocate(sh_final2, enum, .after=id)
```

```{r wrangling2, eval=F, echo=F, cache=T, results=FALSE, warning=FALSE, message=FALSE} 

#Make a new dataset called 'peri'
peri<- sh_final
str(peri)

#drop useless variables 
library(dplyr)

peri2<-peri[!(colnames(peri) %in% c("t.start2", "NotifID", "PtRRTID", "PtARID", "NotifID.x", "PtRRTID.x", "SDPID.x", "givendate.x", "NotifID.y", "PtRRTID.y", "SDPID.y", "PtRenalSpecify", "PtRenal2", "PtRenalSpecify2", "givendate.y"))]

peri2$PtHeight<-replace(peri2$PtHeight, is.na(peri2$PtHeight), 8888)
peri2$PtWeight<-replace(peri2$PtWeight, is.na(peri2$PtWeight), 8888)

peri2[is.na(peri2)]<-0

#23-43
#convert PtRenal1 from long to wide 
plot_bar(peri2$PtRenal1)

## | use the newer pivot_wider() from tidyverse/tidyr to make the conversion 
## | add a new column to indicate primary renal disease is "present" 
peri2['prim.dis']<-as.integer(1)
class(peri2$prim.dis)

peri2 <- peri2 %>%
  pivot_wider(
    names_from = PtRenal1, 
    values_from = prim.dis
  )

##  | rename the new columns 
names(peri2)[48:57] <- c("toxic.nephropthy", "hpt","glomerulonephritis", "dm", "unknown", "obs.uropathy", "others", "ADPKD","nil", "miss")

##  | drop column nil and miss 
peri2<- peri2[,-(56:57)]

peri2[is.na(peri2)]<-0

##  | drop PtHeight and PtWeight (for now since too many missing data)
peri.final <- subset(peri2, select = -c(PtHeight, PtWeight, Modality))

```




##### PRESENTATION TO DR ONG LM ##### 

Overall aim: 
1. Comprehensive analysis of peritonitis risk among PD patients in Malaysia

Specifically to answer the following question: 

1. How long it takes from first PD to developing first peritonitis? 
2. How does this interval fluctuates according to different risk factors? 
3. How does the interval  to first peritonitis and risk change when competing risks (i.e. transfer to HD, death, transplanted, partial recover of kidney functions) are considered?
4. What is the relationship between suffering peritonitis with a terminal event (defined as transfer to HD, and death)

# Kaplan-Meier and Log-rank test 

source: http://www.sthda.com/english/wiki/survival-analysis-basics#fit-complex-survival-curves 

The Kaplan-Meier (KM) survival estimate is a non-parametric approach to estimate the survival probability. The survival function (i.e. the mathematical equation that calculates the survival probability) is defined as: $$S(t_i)=S(t_{i-1})(1-d_i/n_i)$$

where, 

- S(t~i~) is the survival probability at time t~i~
- d~i~ is th number of events at time t~i~
- n~i~ is the number of patients alive just before time t~i~ 

The KM survival curve and the logrank test are *univariate analysis* which are limited to one factor when describing the survival probability. This factor must be a categorical variable. 

KM survival analysis will ignore the impact from all other factors (except the one included in the model). Thus, KM survival analysis might be biased in its estimates. \textcolor{red}{need to find literature that proved KM is biased in its estimates}

## KM survival curves by patient gender 

Based on the data extracted from the NRR (2014-2018, PD-only), the median peritonitis-free survival by gender and the survival curves are shown below: 

```{r, eval=T, echo=F, cache=T, warning=F, message=F, results=T}

print(fit.kaplan.sex)
KM_sex

```

There was no significant difference between both gender (p=0.95) in terms of the duration to first peritonitis (peritonitis-free interval). 

The complement of the KM estimate of the survival function (1-S(t)) provides an estimat of the *cumulative incidence* of events over time. Based on our analysis, $S(t_{60m})=0.473$. 

```{r, eval=T, echo=F, cache=T, warning=F, message=F, results=T}

summary(fit.kaplan, times=60)
```

The cumulative incidence of peritonitis over 60 months was 0.527. 

An interesting observation is the discrepancy between the risk table and the KM survival curve. The risk table suggests that the number of patient at risk of peritonitis decreased substantially but the survival curve remained plateau at approximately 0.5. A high density in the censor plot implied that a large number of our subjects did not develop peritonitis before they were censored. 



# Cox regression analysis 

Source: http://www.sthda.com/english/wiki/cox-proportional-hazards-model#:~:text=The%20Cox%20model%20can%20be,term%20that%20varies%20with%20time. 

We know that the risk of peritonitis is affected by multiple clinical and socio-economical factors (need quotes) simultaneously, but KM analysis could not handle *multivariate approach*. 

The Cox proportional hazard (PH) model is appropriate to evaluate multiple factors that influence survival. The Cox PH model produces *hazard rate*, which is the rate of an event of interest (like peritonitis) happening at a particular time point. This *hazard rate* can be expressed through the *hazard function*, as below: $$h(t)=h_0(t)*exp(b_1x_1 + b_2x_2 + ... + b_px_p)$$
where, 

- *t* is the survival time 
- h(t) is the hazard function
- x~p~ is the set of p covariates
- b~p~ is the coefficient that measures the effect size of a covariate
- h~0~ is the baseline hazard. This is the valuem of hazard if all covariates x~i~ are equal to zero (i.e. exp(0)=1)

The hazard function can be simplified as $$h(t)=h_0(t)*e^(\sum_{i=1}^{n} bx)$$

The Cox model can be written as a **multiple linear regression** of the **logarithm** of the hazard on variables x~i~, with the baseline hazard being an 'intercept' term that varies with time.

## Univariate Cox regression 

We begin by examining the influence of individual hazard of peritonitis with *univariate Cox regression* analysis. 

The variables included were: age, gender, race, require assistant to perform PD, primary renal diseases, and comorbidities 
```{r, eval=T, echo=F, cache=T, warning=F, message=F, results=T}

res <- t(as.data.frame(univ_results, check.names = FALSE))
as.data.frame(res)

```

The inclusion criteria for the subsequent *multivariate Cox regression* was a p-value of <0.1 (instead of 0.05). *Univariate Cox regression* suggested that Age, comorbidity (cancer and CCF) and primary renal disease from diabetes had significant influence over the hazard rate of peritonitis. These variables were included in the following *multivariate Cox regression* analysis 

## Multivariate Cox regression 
```{r, eval=T, echo=F, cache=T, warning=F, message=F, results=T}

summary(multiv_cox)
```

The *multivariate Cox regression* showed that **race, age, diabetes as the primary renal disease**, and *cancer as comorbidity* were significant hazards for peritonitis. Specifically, *bumiputera Sabah* were at a higher hazard to experience peritonitis, with a hazard ratio (compared to *Malay* as baseline) of 2.46 (95% CI: 1.452 - 4.163). Having dibetes as the primary renal disease and cancer as a comorbidity both caused a higher hazard of experiencing peritonitis. The respective hazard ratio was 1.33 (95% CI: 1.074 - 1.651) an 3.71 (95% CI: 1.184 - 11.622). Older age also caused a higher hazard, with a hazard ratio of 1.01 (95% CI 1.003 - 1.019).   

The survival curve was shown below: (\textcolor{red}{probably don't show the curve})
```{r, eval=T, echo=F, cache=T, warning=F, message=F, results=T}

cox_plot
```

The survival plot indicated that the patient cohort did not reach median survival (i.e. survival probabilit of 0.5) by the end of the 5-years follow-up. This is in contrast to the KM survival plot

# Competing risk analysis 

One of the issue we identified earlier was the large amount of censored cases (see Part 1: Kaplan Meier survival probability). Both the KM survival analysis and Cox PH regression model treat any other causes of "failure" - any other clinical incidents that precluded the occurence of peritonitis, such as transfer to HD, death, transplanted, or partial recovery of kidney function, as censored cases. 

Obviously, these patients were not lost to follow-up. They experience events that stopped them from developing peritonitis. So censoring these patients at the occurence of competing risk event have violated several assumptions for survival analysis: the censored cases were independent and noninformative.

Using KM estimates for incidence funcion in the presence of competing risks typically would produce upward biases in the estimation of the incidence function (quote Austin 2016). To better estimates the incidence of peritonitis while taking into consideration of competing risk, the *Cumulative Incidence Function (CIF)* should be used. It is distinct from $1-S(t)$, and can be defined for the *k*th cause as $$CIF_k(t)=Pr(T\le{t},D=k)$$

Where,

- D is a variable denoting the type of event that occured

CIF~k~(t) denotes the probability of experiencing the *k*th event before time *t* and before the occurence of a different type of event. Different from the KM estimates, the sum of CIF estimates of the incidence of each individual outcomes will equal the CIF estimates of the incidence of the composite outcome consisting all of the competing events. (quote Austin 2016) 


## Hazard function regression 

The hazard function (which is a function of time) describes the instantaneous rate of occurence of the event of interest (peritonitis) in subjects who are *still at risk* of the event. The Cox PH regression model relates the hazard function to a set of covariates but it does not relate the covariates directly to the survival times themselves. 

The hazard ratio describes the relative change in the hazard function in response to one-unit change in the covariate. The *regression coefficient* of the Cox model (without the exponential) describe the relative effect of the covariates on the hazard. In the absence of competing risks, the following relationship holds: $$S(t)=S_0(t)^{e(x,\beta)}$$

Thus, the relative effect (i.e. *regressioin coefficient*) of a given covariate on the **hazard of the outcome** is equal to the relative effect of that covariate on the **logarithm of the survival function**. Thus, any inferences made about the effect of a covariate on the hazard function are equivalent about the effect on prognosis or survival. This direct correspondence between **hazard function** and **incidence** allows conclusions about a risk factor (or variable) increases the risk of an event without specifying whether risk denoted the hazard of an event (the **rate of occurence** of people who are still at risk) or the incidence of the event (the **probability** of the occurence of the event). 

In competing risk analysis, 2 different types of hazard functions are of interest: 

- *cause-specific* hazard function 
- *subdistribution* hazard function 

The *subdistribution* hazard model is also known as the *CIF regression model*, which describes the link between the subdistribution hazard and the effect on the incidence of an event. This model allows the direct prediction of the cumulative incidence for an event of interest using the usual relationship between the hazard and the incidence function. The subdistribution hazard model also allows the estimation of the effect of covariates on the cumulative incidence function for the event of interest. 

